---
title: "ADNI Analysis 2: Logistic Regression"
---

#Load Packages

```{r}
library(lmerTest)
library(corrplot)
library(ggplot2)
```

#Read in Dataset
```{r}
#Reading in Data
# MCI_orig<- data.frame(read.table("/Users/shreyarajagopal/Documents/Research/ADNI_4/Subject_Selection/MCI_Only_MCI_AD_Visits_5_March22.csv",header = T, sep = ","))

MCI <-data.frame(read.table("/Users/shreyarajagopal/Documents/Research/ADNI_4/Subject_Selection/MCI_Jun22.csv",header = T, sep = ","))


MCI = MCI[MCI$RID!=2336,] #Removing subjects due to issues with ICV measurements
#MCI = MCI[MCI$RID!=6632,] #Removing subject due to issues with MMSE and CDRSB
```

#Comparing Dataframes with and without occipital

library(dplyr)

MCI_orig <- distinct(MCI_orig)

MCI_occip <- MCI[,c(1:89)]
MCI_occip <- distinct(MCI_occip)

MCI_intersect <- inner_join(MCI_orig, MCI_occip)



#Retain only pre-conversion MCI visits

```{r}
MCI <- MCI[MCI$Diagnosis == "MCI",]
```

#Making Cohort and Gender into factors (treated as categorical variables) + Relevelling:
```{r}
#Relevelling

#Cohort
MCI$Cohort = factor(MCI$Cohort)
cat("\nCurrent Levels:",levels(MCI$Cohort))
MCI$Cohort<- relevel(MCI$Cohort, "MCI_NC") #MCI_NC = 0, MCI_C = 1
cat("\nNew Levels:",levels(MCI$Cohort))

#Gender
MCI$Gender <- factor(MCI$Gender)
cat("\nGender Levels:",levels(MCI$Gender)) #Female = 0, Male = 1

MCI$Age_sc = scale(MCI$Age)
M_Age = mean(MCI$Age,na.rm = TRUE)
SD_Age = sd(MCI$Age, na.rm = TRUE)
```


#Occipital with only Age as a predictor

```{r}
Occip_mod <- lmer(Occipital ~ Age_sc + (1 + Age_sc|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = TRUE, na.action = na.exclude)
summary(Occip_mod)
```
#Dividing coefficients up by Cohort
```{r}
#Coefficients
Occip_rand_eff <- ranef(Occip_mod)[["RID"]]
Occip_rand_eff <- cbind(RID=rownames(Occip_rand_eff),Occip_rand_eff)
#rownames(Occip_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

Occip_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(Occip_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (Occip_rand_eff[i,]$RID == MCI_C_ids[j])
    {Occip_rand_eff[i,]$Cohort = "MCI_C"}
  }}
```

#Make Cohort a factor

```{r}
#Re-leveling non-converters to be level 0 and converters to be level 1
is.factor(Occip_rand_eff$Cohort)
Occip_rand_eff$Cohort <- factor(Occip_rand_eff$Cohort)
levels(Occip_rand_eff$Cohort)
Occip_rand_eff$Cohort<- relevel(Occip_rand_eff$Cohort, "MCI_NC") #Made No Treatment baseline
levels(Occip_rand_eff$Cohort)  #MCI-NC is level 0 and MCI-C is level 1

```

```{r}
#Computing final slopes for each individual (adding back fixed effects)

#Adding back fixed effects to random effects to get final slope for each individual
Occip_rand_eff$Age_sc = Occip_rand_eff$Age_sc + Occip_mod@beta[2]  
Occip_rand_eff$`(Intercept)`= Occip_rand_eff$`(Intercept)`+ Occip_mod@beta[1]

#Plotting Occipital rates of change
#fitted values
MCI$Occipital_fit = fitted(Occip_mod) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_Occipital_Coeff <- Occip_rand_eff[Occip_rand_eff$Cohort == "MCI_NC",]
MCI_C_Occipital_Coeff <- Occip_rand_eff[Occip_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (Age_sc) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Occipital_MCI_C_slope = mean(MCI_C_Occipital_Coeff$Age_sc, na.rm = TRUE)
Occipital_MCI_NC_slope = mean(MCI_NC_Occipital_Coeff$Age_sc, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Occipital_MCI_C_intercept = mean(MCI_C_Occipital_Coeff$`(Intercept)`, na.rm = TRUE)
Occipital_MCI_NC_intercept = mean(MCI_NC_Occipital_Coeff$`(Intercept)`, na.rm = TRUE)

```

#Occipital Plot - WEAKER Age_sc cohort interaction than entorhinal? Still significant
```{r}
ggplot(MCI, aes(x=Age_sc, y=Occipital_fit, group = RID, color = Cohort))+geom_line(size=1.5) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age_sc, y=Occipital, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 2) + 
  geom_abline(slope = Occipital_MCI_C_slope, intercept = Occipital_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = Occipital_MCI_NC_slope, intercept = Occipital_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject") + ylab("Standardized Occipital Volume (SD units)") + ggtitle("Changes in Occipital Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5)) 
```


#
#ICV with only Age as a predictor

```{r}
ICV_mod <- lmer(ICV ~ Age_sc + (1 + Age_sc|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = TRUE, na.action = na.exclude)
summary(ICV_mod)
```

#Dividing coefficients up by Cohort
```{r}
#Coefficients
ICV_rand_eff <- ranef(ICV_mod)[["RID"]]
ICV_rand_eff <- cbind(RID=rownames(ICV_rand_eff),ICV_rand_eff)
#rownames(ICV_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

ICV_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(ICV_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (ICV_rand_eff[i,]$RID == MCI_C_ids[j])
    {ICV_rand_eff[i,]$Cohort = "MCI_C"}
  }}
```

#Make Cohort a factor

```{r}
#Re-leveling non-converters to be level 0 and converters to be level 1
is.factor(ICV_rand_eff$Cohort)
ICV_rand_eff$Cohort <- factor(ICV_rand_eff$Cohort)
levels(ICV_rand_eff$Cohort)
ICV_rand_eff$Cohort<- relevel(ICV_rand_eff$Cohort, "MCI_NC") #Made No Treatment baseline
levels(ICV_rand_eff$Cohort)  #MCI-NC is level 0 and MCI-C is level 1
```
```{r}
#Computing final slopes for each individual (adding back fixed effects)

#Adding back fixed effects to random effects to get final slope for each individual
ICV_rand_eff$Age_sc = ICV_rand_eff$Age_sc + ICV_mod@beta[2]  
ICV_rand_eff$`(Intercept)`= ICV_rand_eff$`(Intercept)`+ ICV_mod@beta[1]

#Plotting ICV rates of change
#fitted values
MCI$ICV_fit = fitted(ICV_mod) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_ICV_Coeff <- ICV_rand_eff[ICV_rand_eff$Cohort == "MCI_NC",]
MCI_C_ICV_Coeff <- ICV_rand_eff[ICV_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (Age_sc) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
ICV_MCI_C_slope = mean(MCI_C_ICV_Coeff$Age_sc, na.rm = TRUE)
ICV_MCI_NC_slope = mean(MCI_NC_ICV_Coeff$Age_sc, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
ICV_MCI_C_intercept = mean(MCI_C_ICV_Coeff$`(Intercept)`, na.rm = TRUE)
ICV_MCI_NC_intercept = mean(MCI_NC_ICV_Coeff$`(Intercept)`, na.rm = TRUE)

```

#ICV Plot - WEAKER Age_sc cohort interaction than entorhinal? Still significant
```{r}
ggplot(MCI, aes(x=Age_sc, y=ICV_fit, group = RID, color = Cohort))+geom_line(size=1.5) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age_sc, y=ICV, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 2) + 
  geom_abline(slope = ICV_MCI_C_slope, intercept = ICV_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = ICV_MCI_NC_slope, intercept = ICV_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject") + ylab("Standardized ICV Volume (SD units)") + ggtitle("Changes in ICV Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5)) + theme_classic() + theme(axis.text=element_text(size=12),
                                                                          axis.title=element_text(size=14,face="bold"))


```

#Hippocampus with only Age as a predictor

```{r}
Hip_mod <- lmer(Hippocampus ~ Age_sc + (1 + Age_sc|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = TRUE, na.action = na.exclude)
summary(Hip_mod)
```
#Dividing coefficients up by Cohort
```{r}
#Coefficients
Hip_rand_eff <- ranef(Hip_mod)[["RID"]]
Hip_rand_eff <- cbind(RID=rownames(Hip_rand_eff),Hip_rand_eff)
#rownames(Hip_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

Hip_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(Hip_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (Hip_rand_eff[i,]$RID == MCI_C_ids[j])
    {Hip_rand_eff[i,]$Cohort = "MCI_C"}
  }}
```

#Make Cohort a factor

```{r}
#Re-leveling non-converters to be level 0 and converters to be level 1
is.factor(Hip_rand_eff$Cohort)
Hip_rand_eff$Cohort <- factor(Hip_rand_eff$Cohort)
levels(Hip_rand_eff$Cohort)
Hip_rand_eff$Cohort<- relevel(Hip_rand_eff$Cohort, "MCI_NC") #Made No Treatment baseline
levels(Hip_rand_eff$Cohort)  #MCI-NC is level 0 and MCI-C is level 1

```

```{r}
#Computing final slopes for each individual (adding back fixed effects)

#Adding back fixed effects to random effects to get final slope for each individual
Hip_rand_eff$Age_sc = Hip_rand_eff$Age_sc + Hip_mod@beta[2]  
Hip_rand_eff$`(Intercept)`= Hip_rand_eff$`(Intercept)`+ Hip_mod@beta[1]

#Plotting Hippocampus rates of change
#fitted values
MCI$Hippocampus_fit = fitted(Hip_mod) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_Hippocampus_Coeff <- Hip_rand_eff[Hip_rand_eff$Cohort == "MCI_NC",]
MCI_C_Hippocampus_Coeff <- Hip_rand_eff[Hip_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (Age_sc) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Hippocampus_MCI_C_slope = mean(MCI_C_Hippocampus_Coeff$Age_sc, na.rm = TRUE)
Hippocampus_MCI_NC_slope = mean(MCI_NC_Hippocampus_Coeff$Age_sc, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Hippocampus_MCI_C_intercept = mean(MCI_C_Hippocampus_Coeff$`(Intercept)`, na.rm = TRUE)
Hippocampus_MCI_NC_intercept = mean(MCI_NC_Hippocampus_Coeff$`(Intercept)`, na.rm = TRUE)

```

#Hippocampus Plot - WEAKER Age_sc cohort interaction than entorhinal? Still significant
```{r}
ggplot(MCI, aes(x=Age_sc, y=Hippocampus_fit, group = RID, color = Cohort))+geom_line(size=1.5) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age_sc, y=Hippocampus, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 2) + 
  geom_abline(slope = Hippocampus_MCI_C_slope, intercept = Hippocampus_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = Hippocampus_MCI_NC_slope, intercept = Hippocampus_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject") + ylab("Standardized Hippocampus Volume (SD units)") + ggtitle("Changes in Hippocampus Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5)) 
#Checking correlations b/w predictors and outcom
```
#Entorhinal Growth Curves with only Age as a predictor
```{r}
ent_mod <- lmer(Entorhinal ~ Age_sc + (1 + Age_sc|RID), data = MCI, control = lmerControl(optimizer="Nelder_Mead"), REML = TRUE,na.action = na.exclude)

summary(ent_mod)
```

#Dividing coefficients up by Cohort
```{r}
#Coefficients
ent_rand_eff <- ranef(ent_mod)[["RID"]]
ent_rand_eff <- cbind(RID=rownames(ent_rand_eff),ent_rand_eff)
#rownames(ent_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

ent_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(ent_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (ent_rand_eff[i,]$RID == MCI_C_ids[j])
    {ent_rand_eff[i,]$Cohort = "MCI_C"}
  }}

```

#Make Cohort a factor

```{r}
#Re-leveling non-converters to be level 0 and converters to be level 1
is.factor(ent_rand_eff$Cohort)
ent_rand_eff$Cohort <- factor(ent_rand_eff$Cohort)
levels(ent_rand_eff$Cohort)
ent_rand_eff$Cohort<- relevel(ent_rand_eff$Cohort, "MCI_NC") #Made No Treatment baseline
levels(ent_rand_eff$Cohort)  #MCI-NC is level 0 and MCI-C is level 1
```

```{r}
#Computing final slopes for each individual (adding back fixed effects)

#Adding back fixed effects to random effects to get final slope for each individual
ent_rand_eff$Age_sc = ent_rand_eff$Age_sc + ent_mod@beta[2]  
ent_rand_eff$`(Intercept)`= ent_rand_eff$`(Intercept)`+ ent_mod@beta[1]

#Plotting Entorhinal rates of change
#fitted values
MCI$Entorhinal_fit = fitted(ent_mod) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_Entorhinal_Coeff <- ent_rand_eff[ent_rand_eff$Cohort == "MCI_NC",]
MCI_C_Entorhinal_Coeff <- ent_rand_eff[ent_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (Age_sc) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Entorhinal_MCI_C_slope = mean(MCI_C_Entorhinal_Coeff$Age_sc, na.rm = TRUE)
Entorhinal_MCI_NC_slope = mean(MCI_NC_Entorhinal_Coeff$Age_sc, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Entorhinal_MCI_C_intercept = mean(MCI_C_Entorhinal_Coeff$`(Intercept)`, na.rm = TRUE)
Entorhinal_MCI_NC_intercept = mean(MCI_NC_Entorhinal_Coeff$`(Intercept)`, na.rm = TRUE)

```

#Entorhinal Plot - WEAKER Age_sc cohort interaction than entorhinal? Still significant
```{r}
ggplot(MCI, aes(x=Age_sc, y=Entorhinal_fit, group = RID, color = Cohort))+geom_line(size=1.5) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age_sc, y=Entorhinal, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 2) + 
  geom_abline(slope = Entorhinal_MCI_C_slope, intercept = Entorhinal_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = Entorhinal_MCI_NC_slope, intercept = Entorhinal_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject") + ylab("Standardized Entorhinal Volume (SD units)") + ggtitle("Changes in Entorhinal Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5)) 
```
```{r}
cormat1 = merge(Occip_rand_eff, ICV_rand_eff,by = "RID")

colnames(cormat1)[2] <- "Occip_intercept"
colnames(cormat1)[3] <- "Occip_slope"
colnames(cormat1)[5] <-"ICV_intercept"
colnames(cormat1)[6] <-"ICV_slope"
colnames(cormat1)[4]<-"Cohort"

cormat2 = merge(Hip_rand_eff, ent_rand_eff,by = "RID")

colnames(cormat2)[2] <- "Hip_intercept"
colnames(cormat2)[3] <- "Hip_slope"
colnames(cormat2)[5] <-"ent_intercept"
colnames(cormat2)[6] <-"ent_slope"
colnames(cormat2)[4]<-"Cohort"


```



#Z-scoring slopes
```{r}



M_Occip_Slope = mean(cormat1$Occip_slope,na.rm = TRUE)
SD_Occip_Slope = sd(cormat1$Occip_slope,na.rm = TRUE)

M_ICV_Slope = mean(cormat1$ICV_slope,na.rm = TRUE)
SD_ICV_Slope = sd(cormat1$ICV_slope,na.rm = TRUE)

cormat1$Occip_slope = scale(cormat1$Occip_slope)
cormat1$ICV_slope = scale(cormat1$ICV_slope)

M_Hip_Slope = mean(cormat2$Hip_slope,na.rm = TRUE)
SD_Hip_Slope = sd(cormat2$Hip_slope,na.rm = TRUE)


M_ent_Slope = mean(cormat2$ent_slope,na.rm = TRUE)
SD_ent_Slope = sd(cormat2$ent_slope,na.rm = TRUE)

cormat2$Hip_slope = scale(cormat2$Hip_slope)
cormat2$ent_slope = scale(cormat2$ent_slope)

cormat = merge(cormat1,cormat2,by = "RID")
```

```{r}
plot(cormat$Occip_slope,cormat$Hip_slope,xlab = "Occipital Slopes", ylab = "Hippocampal Slopes",main = "Random Slopes Correlation",col = ifelse(cormat$Cohort.x == "MCI_C",'blue','red'))
#text(-2.3,0.35,paste("r = ",round(c_Hippocampus_F3$statistic,2), "\np <0.001"))

abline(lm(cormat$Occip_slope~cormat$Hip_slope))

plot(cormat$Occip_slope,cormat$ent_slope,xlab = "Occipital Slopes", ylab = "entorhinal Slopes",main = "Random Slopes Correlation",col = ifelse(cormat$Cohort.x == "MCI_C",'blue','red'))
#text(-2.3,0.35,paste("r = ",round(c_entpocampus_F3$statistic,2), "\np <0.001"))

abline(lm(cormat$Occip_slope~cormat$ent_slope))
```

#Predicting cohort from Occipital Slopes
```{r}
LR_Occip_null <- glm(Cohort ~ 1, data = cormat1, family = binomial)
LR_Occip <- glm(Cohort ~ Occip_slope, data = cormat1, family = binomial) 
anova(LR_Occip_null, LR_Occip)
1 - pchisq(8.4849, 1)
summary(LR_Occip)
summary(LR_Occip_null)
exp(LR_Occip$coefficients)
```

#Predicting cohort from ICVital Slopes
```{r}
LR_ICV_null <- glm(Cohort ~ 1, data = cormat1, family = binomial)
LR_ICV <- glm(Cohort ~ ICV_slope, data = cormat1, family = binomial) 
anova(LR_ICV_null, LR_ICV)
1 - pchisq(8.4849, 1)
summary(LR_ICV)
summary(LR_ICV_null)
exp(LR_ICV$coefficients)
```
#LR Plots
```{r}
library(effects)
plot(Effect(focal.predictors ="Occip_slope", mod = LR_Occip), xlab = "Rate of change of Occipital volume", ylab = "Probability of progression to to Dementia", main = "Predicting Progression to Dementia - Occipital Rates")
plot(Effect(focal.predictors ="ICV_slope", mod = LR_ICV), xlab = "Rate of change of ICV Volume", ylab = "Probability of progression to to Dementia", main = "Predicting Progression to Dementia - ICV Rates")



```

