---
title: "ADNI Analysis 2: Logistic Regression"
---

#Load Packages

```{r}
library(lmerTest)
library(corrplot)
library(ggplot2)
```

#Read in Dataset
```{r}
#Reading in Data
# MCI_orig<- data.frame(read.table("/Users/shreyarajagopal/Documents/Research/ADNI_4/Subject_Selection/MCI_Only_MCI_AD_Visits_5_March22.csv",header = T, sep = ","))

MCI <-data.frame(read.table("/Users/shreyarajagopal/Documents/Research/ADNI_4/Subject_Selection/MCI_Jun22.csv",header = T, sep = ","))


MCI = MCI[MCI$RID!=2336,] #Removing subjects due to issues with ICV measurements
#MCI = MCI[MCI$RID!=6632,] #Removing subject due to issues with MMSE and CDRSB
```

#Comparing Dataframes with and without occipital

library(dplyr)

MCI_orig <- distinct(MCI_orig)

MCI_occip <- MCI[,c(1:89)]
MCI_occip <- distinct(MCI_occip)

MCI_intersect <- inner_join(MCI_orig, MCI_occip)



#Retain only pre-conversion MCI visits

```{r}
MCI <- MCI[MCI$Diagnosis == "MCI",]
```

#Making Cohort and Gender into factors (treated as categorical variables) + Relevelling:
```{r}
#Relevelling

#Cohort
MCI$Cohort = factor(MCI$Cohort)
cat("\nCurrent Levels:",levels(MCI$Cohort))
MCI$Cohort<- relevel(MCI$Cohort, "MCI_NC") #MCI_NC = 0, MCI_C = 1
cat("\nNew Levels:",levels(MCI$Cohort))

#Gender
MCI$Gender <- factor(MCI$Gender)
cat("\nGender Levels:",levels(MCI$Gender)) #Female = 0, Male = 1

```

#Age of Onset
```{r}
#install.packages("dplyr")
library(dplyr)

# Assuming MCI is your original dataframe
new_dataframe <- MCI %>%
  #filter(Dx_bl == "CN") %>%
  group_by(RID) %>%
  arrange(Visit, .by_group = TRUE) %>%
  slice(1) %>%
  select(RID, Age, Visit, Dx_bl, Diagnosis,Cohort)
```


#Standardizing Hippocampus and EC volume wrt ICV

```{r}
MCI$Hippocampus_icv_absolute = MCI$Hippocampus/MCI$ICV
MCI$Entorhinal_icv_absolute = MCI$Entorhinal/MCI$ICV
MCI$Fusiform_icv_absolute = MCI$Fusiform/MCI$ICV
MCI$Ventricles_icv_absolute = MCI$Ventricles/MCI$ICV
MCI$Occipital_icv_absolute = MCI$Occipital/MCI$ICV
MCI$WholeBrain_icv_absolute = MCI$WholeBrain/MCI$ICV

#Storing SD and means of normalized Regional brain volumes
M_Hippocampus_icv = mean(MCI$Hippocampus_icv_absolute,na.rm = TRUE)
SD_Hippocampus_icv = sd(MCI$Hippocampus_icv_absolute,na.rm = TRUE)

M_Entorhinal_icv = mean(MCI$Entorhinal_icv_absolute,na.rm = TRUE)
S_Entorhinal_icv = sd(MCI$Entorhinal_icv_absolute,na.rm = TRUE)

M_Fusiform_icv = mean(MCI$Fusiform_icv_absolute,na.rm = TRUE)
SD_Fusiform_icv = sd(MCI$Fusiform_icv_absolute,na.rm = TRUE)

M_Occipital_icv = mean(MCI$Occipital_icv_absolute,na.rm = TRUE)
SD_Occipital_icv = sd(MCI$Occipital_icv_absolute,na.rm = TRUE)

M_Ventricles_icv = mean(MCI$Ventricles_icv_absolute,na.rm = TRUE)
SD_Ventricles_icv = sd(MCI$Ventricles_icv_absolute,na.rm = TRUE)

M_MWholeBrain_icv = mean(MCI$WholeBrain_icv_absolute,na.rm = TRUE)
SD_WholeBrain_icv = sd(MCI$WholeBrain_icv_absolute,na.rm = TRUE)

#Z-scoring variables
MCI$Hippocampus_icv = scale(MCI$Hippocampus_icv_absolute)
MCI$Entorhinal_icv = scale(MCI$Entorhinal_icv_absolute)
MCI$Fusiform_icv = scale(MCI$Fusiform_icv_absolute)
MCI$Ventricles_icv = scale(MCI$Ventricles_icv_absolute)
MCI$Occipital_icv = scale(MCI$Occipital_icv_absolute)
MCI$WholeBrain_icv = scale(MCI$WholeBrain_icv_absolute)

MCI$Age_sc = scale(MCI$Age)
M_Age = mean(MCI$Age,na.rm = TRUE)
SD_Age = sd(MCI$Age, na.rm = TRUE)
```

#Hippocampus with only Age as a predictor

```{r}
Hip_mod <- lmer(Hippocampus_icv ~ Age_sc + (1 + Age_sc|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = TRUE, na.action = na.exclude)
summary(Hip_mod)
```

#Dividing coefficients up by Cohort
```{r}
#Coefficients
Hip_rand_eff <- ranef(Hip_mod)[["RID"]]
Hip_rand_eff <- cbind(RID=rownames(Hip_rand_eff),Hip_rand_eff)
#rownames(Hip_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

Hip_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(Hip_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (Hip_rand_eff[i,]$RID == MCI_C_ids[j])
    {Hip_rand_eff[i,]$Cohort = "MCI_C"}
  }}
```

#Make Cohort a factor

```{r}
#Re-leveling non-converters to be level 0 and converters to be level 1
is.factor(Hip_rand_eff$Cohort)
Hip_rand_eff$Cohort <- factor(Hip_rand_eff$Cohort)
levels(Hip_rand_eff$Cohort)
Hip_rand_eff$Cohort<- relevel(Hip_rand_eff$Cohort, "MCI_NC") #Made No Treatment baseline
levels(Hip_rand_eff$Cohort)  #MCI-NC is level 0 and MCI-C is level 1
```

```{r}
#Computing final slopes for each individual (adding back fixed effects)

#Adding back fixed effects to random effects to get final slope for each individual
Hip_rand_eff$Age_sc = Hip_rand_eff$Age_sc + Hip_mod@beta[2]  
Hip_rand_eff$`(Intercept)`= Hip_rand_eff$`(Intercept)`+ Hip_mod@beta[1]

#Plotting Hippocampus rates of change
#fitted values
MCI$Hippocampus_fit = fitted(Hip_mod) #use na.exclude in the model, otherwise size of orignial data and fitted data will be differHip

#Dividing up coefficiHips by cohort
MCI_NC_Hippocampus_Coeff <- Hip_rand_eff[Hip_rand_eff$Cohort == "MCI_NC",]
MCI_C_Hippocampus_Coeff <- Hip_rand_eff[Hip_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (age) coefficiHip for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Hippocampus_MCI_C_slope = mean(MCI_C_Hippocampus_Coeff$Age_sc, na.rm = TRUE)
Hippocampus_MCI_NC_slope = mean(MCI_NC_Hippocampus_Coeff$Age_sc, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Hippocampus_MCI_C_intercept = mean(MCI_C_Hippocampus_Coeff$`(Intercept)`, na.rm = TRUE)
Hippocampus_MCI_NC_intercept = mean(MCI_NC_Hippocampus_Coeff$`(Intercept)`, na.rm = TRUE)

```

#Hippocampus Plot - WEAKER age cohort interaction than Hiporhinal? Still significant
```{r}
ggplot(MCI, aes(x=Age_sc, y=Hippocampus_fit, group = RID, color = Cohort))+geom_line(size=1.5) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age_sc, y=Hippocampus_icv, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 2) + 
  geom_abline(slope = Hippocampus_MCI_C_slope, intercept = Hippocampus_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = Hippocampus_MCI_NC_slope, intercept = Hippocampus_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject") + ylab("Standardized Hippocampus Volume (SD units)") + ggtitle("Changes in Hippocampus Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5)) 
```




#Entorhinal Growth Curves with only Age as a predictor
```{r}
ent_mod <- lmer(Entorhinal_icv ~ Age_sc + (1 + Age_sc|RID), data = MCI, control = lmerControl(optimizer="Nelder_Mead"), REML = TRUE,na.action = na.exclude)

summary(ent_mod)
```

#Dividing coefficients up by Cohort
```{r}
#Coefficients
ent_rand_eff <- ranef(ent_mod)[["RID"]]
ent_rand_eff <- cbind(RID=rownames(ent_rand_eff),ent_rand_eff)
#rownames(ent_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

ent_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(ent_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (ent_rand_eff[i,]$RID == MCI_C_ids[j])
    {ent_rand_eff[i,]$Cohort = "MCI_C"}
  }}

```

#Make Cohort a factor

```{r}
#Re-leveling non-converters to be level 0 and converters to be level 1
is.factor(ent_rand_eff$Cohort)
ent_rand_eff$Cohort <- factor(ent_rand_eff$Cohort)
levels(ent_rand_eff$Cohort)
ent_rand_eff$Cohort<- relevel(ent_rand_eff$Cohort, "MCI_NC") #Made No Treatment baseline
levels(ent_rand_eff$Cohort)  #MCI-NC is level 0 and MCI-C is level 1
```

```{r}
#Computing final slopes for each individual (adding back fixed effects)

#Adding back fixed effects to random effects to get final slope for each individual
ent_rand_eff$Age_sc = ent_rand_eff$Age_sc + ent_mod@beta[2]  
ent_rand_eff$`(Intercept)`= ent_rand_eff$`(Intercept)`+ ent_mod@beta[1]

#Plotting Entorhinal rates of change
#fitted values
MCI$Entorhinal_fit = fitted(ent_mod) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_Entorhinal_Coeff <- ent_rand_eff[ent_rand_eff$Cohort == "MCI_NC",]
MCI_C_Entorhinal_Coeff <- ent_rand_eff[ent_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (Age_sc) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Entorhinal_MCI_C_slope = mean(MCI_C_Entorhinal_Coeff$Age_sc, na.rm = TRUE)
Entorhinal_MCI_NC_slope = mean(MCI_NC_Entorhinal_Coeff$Age_sc, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Entorhinal_MCI_C_intercept = mean(MCI_C_Entorhinal_Coeff$`(Intercept)`, na.rm = TRUE)
Entorhinal_MCI_NC_intercept = mean(MCI_NC_Entorhinal_Coeff$`(Intercept)`, na.rm = TRUE)

```

#Entorhinal Plot - WEAKER Age_sc cohort interaction than entorhinal? Still significant
```{r}
ggplot(MCI, aes(x=Age_sc, y=Entorhinal_fit, group = RID, color = Cohort))+geom_line(size=1.5) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age_sc, y=Entorhinal_icv, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 2) + 
  geom_abline(slope = Entorhinal_MCI_C_slope, intercept = Entorhinal_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = Entorhinal_MCI_NC_slope, intercept = Entorhinal_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject") + ylab("Standardized Entorhinal Volume (SD units)") + ggtitle("Changes in Entorhinal Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5)) 
```

#Fusiform with only Age as a predictor

```{r}
Fus_mod <- lmer(Fusiform_icv ~ Age_sc + (1 + Age_sc|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = TRUE, na.action = na.exclude)
summary(Fus_mod)
```
#Dividing coefficients up by Cohort
```{r}
#Coefficients
Fus_rand_eff <- ranef(Fus_mod)[["RID"]]
Fus_rand_eff <- cbind(RID=rownames(Fus_rand_eff),Fus_rand_eff)
#rownames(Fus_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

Fus_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(Fus_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (Fus_rand_eff[i,]$RID == MCI_C_ids[j])
    {Fus_rand_eff[i,]$Cohort = "MCI_C"}
  }}
```

#Make Cohort a factor

```{r}
#Re-leveling non-converters to be level 0 and converters to be level 1
is.factor(Fus_rand_eff$Cohort)
Fus_rand_eff$Cohort <- factor(Fus_rand_eff$Cohort)
levels(Fus_rand_eff$Cohort)
Fus_rand_eff$Cohort<- relevel(Fus_rand_eff$Cohort, "MCI_NC") #Made No Treatment baseline
levels(Fus_rand_eff$Cohort)  #MCI-NC is level 0 and MCI-C is level 1
```


```{r}
#Computing final slopes for each individual (adding back fixed effects)

#Adding back fixed effects to random effects to get final slope for each individual
Fus_rand_eff$Age_sc = Fus_rand_eff$Age_sc + Fus_mod@beta[2]  
Fus_rand_eff$`(Intercept)`= Fus_rand_eff$`(Intercept)`+ Fus_mod@beta[1]

#Plotting Fusiform rates of change
#fitted values
MCI$Fusiform_fit = fitted(Fus_mod) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_Fusiform_Coeff <- Fus_rand_eff[Fus_rand_eff$Cohort == "MCI_NC",]
MCI_C_Fusiform_Coeff <- Fus_rand_eff[Fus_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (Age_sc) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Fusiform_MCI_C_slope = mean(MCI_C_Fusiform_Coeff$Age_sc, na.rm = TRUE)
Fusiform_MCI_NC_slope = mean(MCI_NC_Fusiform_Coeff$Age_sc, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Fusiform_MCI_C_intercept = mean(MCI_C_Fusiform_Coeff$`(Intercept)`, na.rm = TRUE)
Fusiform_MCI_NC_intercept = mean(MCI_NC_Fusiform_Coeff$`(Intercept)`, na.rm = TRUE)

```

#Fusiform Plot - WEAKER Age_sc cohort interaction than entorhinal? Still significant
```{r}
ggplot(MCI, aes(x=Age_sc, y=Fusiform_fit, group = RID, color = Cohort))+geom_line(size=1.5) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age_sc, y=Fusiform_icv, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 2) + 
  geom_abline(slope = Fusiform_MCI_C_slope, intercept = Fusiform_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = Fusiform_MCI_NC_slope, intercept = Fusiform_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject") + ylab("Standardized Fusiform Volume (SD units)") + ggtitle("Changes in Fusiform Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5)) 
```




#Ventricles with only Age as a predictor

```{r}
Ven_mod <- lmer(Ventricles_icv ~ Age_sc + (1 + Age_sc|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = TRUE, na.action = na.exclude)
summary(Ven_mod)
```
#Dividing coefficients up by Cohort
```{r}
#Coefficients
Ven_rand_eff <- ranef(Ven_mod)[["RID"]]
Ven_rand_eff <- cbind(RID=rownames(Ven_rand_eff),Ven_rand_eff)
#rownames(Ven_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

Ven_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(Ven_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (Ven_rand_eff[i,]$RID == MCI_C_ids[j])
    {Ven_rand_eff[i,]$Cohort = "MCI_C"}
  }}
```

#Make Cohort a factor

```{r}
#Re-leveling non-converters to be level 0 and converters to be level 1
is.factor(Ven_rand_eff$Cohort)
Ven_rand_eff$Cohort <- factor(Ven_rand_eff$Cohort)
levels(Ven_rand_eff$Cohort)
Ven_rand_eff$Cohort<- relevel(Ven_rand_eff$Cohort, "MCI_NC") #Made No Treatment baseline
levels(Ven_rand_eff$Cohort)  #MCI-NC is level 0 and MCI-C is level 1
```
```{r}
#Computing final slopes for each individual (adding back fixed effects)

#Adding back fixed effects to random effects to get final slope for each individual
Ven_rand_eff$Age_sc = Ven_rand_eff$Age_sc + Ven_mod@beta[2]  
Ven_rand_eff$`(Intercept)`= Ven_rand_eff$`(Intercept)`+ Ven_mod@beta[1]

#Plotting Ventricle rates of change
#fitted values
MCI$Ventricle_fit = fitted(Ven_mod) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_Ventricle_Coeff <- Ven_rand_eff[Ven_rand_eff$Cohort == "MCI_NC",]
MCI_C_Ventricle_Coeff <- Ven_rand_eff[Ven_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (Age_sc) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Ventricle_MCI_C_slope = mean(MCI_C_Ventricle_Coeff$Age_sc, na.rm = TRUE)
Ventricle_MCI_NC_slope = mean(MCI_NC_Ventricle_Coeff$Age_sc, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Ventricle_MCI_C_intercept = mean(MCI_C_Ventricle_Coeff$`(Intercept)`, na.rm = TRUE)
Ventricle_MCI_NC_intercept = mean(MCI_NC_Ventricle_Coeff$`(Intercept)`, na.rm = TRUE)

```

#Ventricle Plot - WEAKER Age_sc cohort interaction than entorhinal? Still significant
```{r}
ggplot(MCI, aes(x=Age_sc, y=Ventricle_fit, group = RID, color = Cohort))+geom_line(size=1.5) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age_sc, y=Ventricles_icv, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 2) + 
  geom_abline(slope = Ventricle_MCI_C_slope, intercept = Ventricle_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = Ventricle_MCI_NC_slope, intercept = Ventricle_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject") + ylab("Standardized Ventricle Volume (SD units)") + ggtitle("Changes in Ventricle Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5)) 
```
#Occipital with only Age as a predictor

```{r}
Occip_mod <- lmer(Occipital_icv ~ Age_sc + (1 + Age_sc|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = TRUE, na.action = na.exclude)
summary(Occip_mod)
```
#Dividing coefficients up by Cohort
```{r}
#Coefficients
Occip_rand_eff <- ranef(Occip_mod)[["RID"]]
Occip_rand_eff <- cbind(RID=rownames(Occip_rand_eff),Occip_rand_eff)
#rownames(Occip_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

Occip_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(Occip_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (Occip_rand_eff[i,]$RID == MCI_C_ids[j])
    {Occip_rand_eff[i,]$Cohort = "MCI_C"}
  }}
```

#Make Cohort a factor

```{r}
#Re-leveling non-converters to be level 0 and converters to be level 1
is.factor(Occip_rand_eff$Cohort)
Occip_rand_eff$Cohort <- factor(Occip_rand_eff$Cohort)
levels(Occip_rand_eff$Cohort)
Occip_rand_eff$Cohort<- relevel(Occip_rand_eff$Cohort, "MCI_NC") #Made No Treatment baseline
levels(Occip_rand_eff$Cohort)  #MCI-NC is level 0 and MCI-C is level 1

```

```{r}
#Computing final slopes for each individual (adding back fixed effects)

#Adding back fixed effects to random effects to get final slope for each individual
Occip_rand_eff$Age_sc = Occip_rand_eff$Age_sc + Occip_mod@beta[2]  
Occip_rand_eff$`(Intercept)`= Occip_rand_eff$`(Intercept)`+ Occip_mod@beta[1]

#Plotting Occipital rates of change
#fitted values
MCI$Occipital_fit = fitted(Occip_mod) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_Occipital_Coeff <- Occip_rand_eff[Occip_rand_eff$Cohort == "MCI_NC",]
MCI_C_Occipital_Coeff <- Occip_rand_eff[Occip_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (Age_sc) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Occipital_MCI_C_slope = mean(MCI_C_Occipital_Coeff$Age_sc, na.rm = TRUE)
Occipital_MCI_NC_slope = mean(MCI_NC_Occipital_Coeff$Age_sc, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Occipital_MCI_C_intercept = mean(MCI_C_Occipital_Coeff$`(Intercept)`, na.rm = TRUE)
Occipital_MCI_NC_intercept = mean(MCI_NC_Occipital_Coeff$`(Intercept)`, na.rm = TRUE)

```

#Occipital Plot - WEAKER Age_sc cohort interaction than entorhinal? Still significant
```{r}
ggplot(MCI, aes(x=Age_sc, y=Occipital_fit, group = RID, color = Cohort))+geom_line(size=1.5) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age_sc, y=Occipital_icv, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 2) + 
  geom_abline(slope = Occipital_MCI_C_slope, intercept = Occipital_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = Occipital_MCI_NC_slope, intercept = Occipital_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject") + ylab("Standardized Occipital Volume (SD units)") + ggtitle("Changes in Occipital Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5)) 
```


#WholeBrain with only Age as a predictor

```{r}
WB_mod <- lmer(WholeBrain_icv ~ Age_sc + (1 + Age_sc|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = TRUE, na.action = na.exclude)
summary(WB_mod)
```

#Dividing coefficients up by Cohort
```{r}
#Coefficients
WB_rand_eff <- ranef(WB_mod)[["RID"]]
WB_rand_eff <- cbind(RID=rownames(WB_rand_eff),WB_rand_eff)
#rownames(WB_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

WB_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(WB_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (WB_rand_eff[i,]$RID == MCI_C_ids[j])
    {WB_rand_eff[i,]$Cohort = "MCI_C"}
  }}
```

#Make Cohort a factor

```{r}
#Re-leveling non-converters to be level 0 and converters to be level 1
is.factor(WB_rand_eff$Cohort)
WB_rand_eff$Cohort <- factor(WB_rand_eff$Cohort)
levels(WB_rand_eff$Cohort)
WB_rand_eff$Cohort<- relevel(WB_rand_eff$Cohort, "MCI_NC") #Made No Treatment baseline
levels(WB_rand_eff$Cohort)  #MCI-NC is level 0 and MCI-C is level 1
```
```{r}
#Computing final slopes for each individual (adding back fixed effects)

#Adding back fixed effects to random effects to get final slope for each individual
WB_rand_eff$Age_sc = WB_rand_eff$Age_sc + WB_mod@beta[2]  
WB_rand_eff$`(Intercept)`= WB_rand_eff$`(Intercept)`+ WB_mod@beta[1]

#Plotting WholeBrain rates of change
#fitted values
MCI$WholeBrain_fit = fitted(WB_mod) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_WholeBrain_Coeff <- WB_rand_eff[WB_rand_eff$Cohort == "MCI_NC",]
MCI_C_WholeBrain_Coeff <- WB_rand_eff[WB_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (Age_sc) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
WholeBrain_MCI_C_slope = mean(MCI_C_WholeBrain_Coeff$Age_sc, na.rm = TRUE)
WholeBrain_MCI_NC_slope = mean(MCI_NC_WholeBrain_Coeff$Age_sc, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
WholeBrain_MCI_C_intercept = mean(MCI_C_WholeBrain_Coeff$`(Intercept)`, na.rm = TRUE)
WholeBrain_MCI_NC_intercept = mean(MCI_NC_WholeBrain_Coeff$`(Intercept)`, na.rm = TRUE)

```

#WholeBrain Plot - WEAKER Age_sc cohort interaction than entorhinal? Still significant
```{r}
ggplot(MCI, aes(x=Age_sc, y=WholeBrain_fit, group = RID, color = Cohort))+geom_line(size=1.5) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age_sc, y=WholeBrain_icv, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 2) + 
  geom_abline(slope = WholeBrain_MCI_C_slope, intercept = WholeBrain_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = WholeBrain_MCI_NC_slope, intercept = WholeBrain_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject") + ylab("Standardized WholeBrain Volume (SD units)") + ggtitle("Changes in WholeBrain Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5)) + theme_classic() + theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))


```
#Age of onset and slope relations
```{r}
library(dplyr)

new_dataframe$RID <- as.character(new_dataframe$RID)
# Assuming new_dataframe_2 and Hip_rand_eff are your dataframes
#Hip_rand_eff_updated <- left_join(Hip_rand_eff, new_dataframe %>% select(RID, Age), by = "RID")
model_hip <- lm(Age ~ Age_sc, data = Hip_rand_eff_updated)

# Summary of the regression model
summary(model_hip)

# Assuming new_dataframe_2 and Hip_rand_eff are your dataframes
#Ent_rand_eff_updated <- left_join(ent_rand_eff, new_dataframe %>% select(RID, Age), by = "RID")
model_ent <- lm(Age ~ Age_sc, data = Ent_rand_eff_updated)
# Summary of the regression model
summary(model_ent)


# Assuming new_dataframe_2 and Hip_rand_eff are your dataframes
#Ven_rand_eff_updated <- left_join(Ven_rand_eff, new_dataframe %>% select(RID, Age), by = "RID")
model_ven <- lm(Age ~ Age_sc, data = Ven_rand_eff_updated)
# Summary of the regression model
summary(model_ven)

# Assuming new_dataframe_2 and Hip_rand_eff are your dataframes
WB_rand_eff_updated <- left_join(WB_rand_eff, new_dataframe %>% select(RID, Age), by = "RID")
model_WB <- lm(Age ~ Age_sc, data = WB_rand_eff_updated)
# Summary of the regression model
summary(model_WB)

# Assuming new_dataframe_2 and Hip_rand_eff are your dataframes
Occip_rand_eff_updated <- left_join(Occip_rand_eff, new_dataframe %>% select(RID, Age), by = "RID")
model_Occip <- lm(Age ~ Age_sc, data = Occip_rand_eff_updated)
# Summary of the regression model
summary(model_Occip)

# Assuming new_dataframe_2 and Hip_rand_eff are your dataframes
Fus_rand_eff_updated <- left_join(Fus_rand_eff, new_dataframe %>% select(RID, Age), by = "RID")
model_fus<- lm(Age ~ Age_sc, data = Fus_rand_eff_updated)
# Summary of the regression model
summary(model_fus)

```





#Checking correlations b/w predictors and outcome

```{r}
#Extracting and renaming required columns
cormat1 = merge(Hip_rand_eff, ent_rand_eff,by = "RID")

colnames(cormat1)[2] <- "Hip_intercept"
colnames(cormat1)[3] <- "Hip_slope"
colnames(cormat1)[5] <-"Ent_intercept"
colnames(cormat1)[6] <-"Ent_slope"
colnames(cormat1)[4]<-"Cohort"

cormat2 = merge(Fus_rand_eff, Ven_rand_eff, by = "RID")

colnames(cormat2)[2] <- "Fus_intercept"
colnames(cormat2)[3] <- "Fus_slope"
colnames(cormat2)[5] <-"Ven_intercept"
colnames(cormat2)[6] <-"Ven_slope"
colnames(cormat2)[4]<-"Cohort"

cormat3 = merge(Occip_rand_eff, WB_rand_eff, by = "RID")
colnames(cormat3)[2] <- "Occip_intercept"
colnames(cormat3)[3] <- "Occip_slope"
colnames(cormat3)[5] <-"WB_intercept"
colnames(cormat3)[6] <-"WB_slope"
colnames(cormat3)[4]<-"Cohort"

#COME BACK HERE JUN 26

cormat4 = merge(cormat1,cormat2,by = "RID")
cormat = merge(cormat4,cormat3,by = "RID")

cormat <- cormat[,c(2,3,5,6,8,9,11,12,14,15,17,18,19)]

colnames(cormat)[13]<-"Cohort"
cormat$Cohort <- factor(cormat$Cohort, levels = c("MCI_NC","MCI_C"))

cormat_corr <- cormat[,c(2,4,6,8,10,12,13)] #Only slopes
cormat_corr$Cohort <- as.numeric(cormat_corr$Cohort)

cormat_corr1 <- cormat
cormat_corr1$Cohort <- as.numeric(cormat_corr1$Cohort)

M = cor(cormat_corr)
corrplot(M, method = 'circle', type = 'lower', insig='blank',addCoef.col ='black', number.cex = 0.8) #if no circle then correlation not significant

K = cor(cormat_corr1)
corrplot(K, method = 'circle', type = 'lower', insig='blank',addCoef.col ='black', number.cex = 0.8) #if no circle then correlation not significant


```
*Interpretation:*

As there is evidence for suppression - EC_slopes significantly correlated with Hip_Slopes but not with outcome, while Hip_slopes significantly correlated with outcome - we use two different logistic regression models for EC_slopes and Hip_slopes

#Z-scoring slopes
```{r}

original_cormat <-cormat

M_Hip_Slope = mean(cormat$Hip_slope,na.rm = TRUE)
SD_Hip_Slope = sd(cormat$Hip_slope,na.rm = TRUE)

M_Ent_Slope = mean(cormat$Ent_slope,na.rm = TRUE)
SD_Ent_Slope = sd(cormat$Ent_slope,na.rm = TRUE)

M_Fus_Slope = mean(cormat$Fus_slope,na.rm = TRUE)
SD_Fus_Slope = sd(cormat$Fus_slope,na.rm = TRUE)

M_Ven_Slope = mean(cormat$Ven_slope,na.rm = TRUE)
SD_Ven_Slope = sd(cormat$Ven_slope,na.rm = TRUE)

M_Occip_Slope = mean(cormat$Occip_slope,na.rm = TRUE)
SD_Occip_Slope = sd(cormat$Occip_slope,na.rm = TRUE)

M_WB_Slope = mean(cormat$WB_slope,na.rm = TRUE)
SD_WB_Slope = sd(cormat$WB_slope,na.rm = TRUE)

cormat$Hip_slope = scale(cormat$Hip_slope)

cormat$Ent_slope = scale(cormat$Ent_slope)

cormat$Fus_slope = scale(cormat$Fus_slope)

cormat$Ven_slope = scale(cormat$Ven_slope)

cormat$Occip_slope = scale(cormat$Occip_slope)

cormat$WB_slope = scale(cormat$WB_slope)




```


```{r}
cormat$Hip_Ven_Ratio = cormat$Hip_slope/cormat$Ven_slope
```
#Logistic Regression predicting cohort from WB_slopes

```{r}
LR_WB_null <-glm(Cohort ~ 1, data = cormat, family = binomial) 
LR_WB <- glm(Cohort ~ WB_slope, data = cormat, family = binomial) #predicting cohort from regression coefficient associated with age 
summary(LR_WB)
summary(LR_WB_null)
anova(LR_WB_null,LR_WB)
1 - pchisq(3.65, 1)
exp(LR_WB$coefficients)
```
#Logistic Regression predicting cohort from Hip_slopes

```{r}
LR_hip_null <-glm(Cohort ~ 1, data = cormat, family = binomial) 
LR_hip <- glm(Cohort ~ Hip_slope, data = cormat, family = binomial) #predicting cohort from regression coefficient associated with age 
summary(LR_hip)
summary(LR_hip_null)
anova(LR_hip_null,LR_hip)
1 - pchisq(27.167, 1)
exp(LR_hip$coefficients)
```

#Logistic Regression predicting cohort from Ent_slopes
```{r}
LR_ent_null <- glm(Cohort ~ 1, data = cormat, family = binomial)
LR_ent <- glm(Cohort ~ Ent_slope, data = cormat, family = binomial) #predicting cohort from regression coefficient associated with age 
anova(LR_ent_null, LR_ent)
1 - pchisq(0.60, 1)
summary(LR_ent)
summary(LR_ent_null)
exp(LR_ent$coefficients)
```

#Logistic Regression predicting cohort from Fus_slopes
```{r}
LR_Fus_null <- glm(Cohort ~ 1, data = cormat, family = binomial)
LR_Fus <- glm(Cohort ~ Fus_slope, data = cormat, family = binomial) #predicting cohort from regression coefficiFus associated with age 
anova(LR_Fus_null, LR_Fus)
1 - pchisq(0.294, 1)
summary(LR_Fus)
summary(LR_Fus_null)
exp(LR_Fus$coefficients)
```

#Logistic Regression predicting cohort from Ven_slopes
```{r}
LR_Ven_null <- glm(Cohort ~ 1, data = cormat, family = binomial)
LR_Ven <- glm(Cohort ~ Ven_slope, data = cormat, family = binomial) #predicting cohort from regression coefficiVen associated with age 
anova(LR_Ven_null, LR_Ven)
1 - pchisq(28.93, 1)
summary(LR_Ven)
summary(LR_Ven_null)
exp(LR_Ven$coefficients)
```

#Predicting cohort from Occipital Slopes
```{r}
LR_Occip_null <- glm(Cohort ~ 1, data = cormat, family = binomial)
LR_Occip <- glm(Cohort ~ Occip_slope, data = cormat, family = binomial) 
anova(LR_Occip_null, LR_Occip)
1 - pchisq(8.4849, 1)
summary(LR_Occip)
summary(LR_Occip_null)
exp(LR_Occip$coefficients)
```
#Predicting cohort from Ratio of Hippocampal/Ventricle Slopes
```{r}
LR_Hip_Ven_null <- glm(Cohort ~ 1, data = cormat, family = binomial)
LR_Hip_Ven <- glm(Cohort ~ Hip_Ven_Ratio, data = cormat, family = binomial) 

summary(LR_Hip_Ven_null)
summary(LR_Hip_Ven)
exp(LR_Hip_Ven)
```

#LR Plots
```{r}
library(effects)
plot(Effect(focal.predictors ="Hip_slope", mod = LR_hip), xlab = "Rate of change of Hippocampus volume", ylab = "Probability of progression to Dementia", main = "Predicting Progression to Dementia - Hippocampal Rates")
plot(Effect(focal.predictors ="Ent_slope", mod = LR_ent), xlab = "Rate of change of EC volume", ylab = "Probability of progression to to Dementia", main = "Predicting Progression to Dementia - EC Rates")
plot(Effect(focal.predictors ="Fus_slope", mod = LR_Fus), xlab = "Rate of change of Fusiform volume", ylab = "Probability of progression to to Dementia", main = "Predicting Progression to Dementia - Fusiform Rates")
plot(Effect(focal.predictors ="Ven_slope", mod = LR_Ven), xlab = "Rate of change of Ventricle volume", ylab = "Probability of progression to to Dementia", main = "Predicting Progression to Dementia - Ventricle Rates")
plot(Effect(focal.predictors ="Occip_slope", mod = LR_Occip), xlab = "Rate of change of Occipital volume", ylab = "Probability of progression to to Dementia", main = "Predicting Progression to Dementia - Occipital Rates")
plot(Effect(focal.predictors ="WB_slope", mod = LR_WB), xlab = "Rate of change of Whole Brain Volume", ylab = "Probability of progression to to Dementia", main = "Predicting Progression to Dementia - Whole Brain Rates")


```

