---
title: "ADNI Analysis 1 : Fitting Growth Curves"
---
Redaing in Data:
```{r}
#Reading in Data
#MCI<- data.frame(read.table("/Users/shreyarajagopal/Documents/Research/ADNI_4/Subject_Selection/MCI_Only_MCI_AD_Visits_5_March22.csv",header = T, sep = ","))

#Dataset with occipital volume
MCI <-data.frame(read.table("/Users/shreyarajagopal/Documents/Research/ADNI_4/Subject_Selection/MCI_Jun22.csv",header = T, sep = ","))
#Number of subjects
cat("MCI_C Number of Subjects: ",length(unique(MCI$RID[MCI$Cohort == "MCI_C"]))) 
cat("\nMCI_NC Number of Subjects: ",length(unique(MCI$RID[MCI$Cohort == "MCI_NC"]))) 
```

```{r}
MCI = MCI[MCI$RID!=2336,] #Removing subjects due to issues with ICV measurements
#MCI = MCI[MCI$RID!=6632,] #Removing subject due to issues with MMSE=22 and CDRSB = 5
#MCI = MCI[MCI$RID!=6341,] #Removing subject due to issues with MMSE = 19
```
#Retain only pre-conversion MCI visits

```{r}
MCI <- MCI[MCI$Diagnosis == "MCI",]
```
#Computing number of repeat visits:
```{r}
library(dplyr)

summary_df <- MCI %>%
  arrange(RID, Visit) %>% # Ensure data is ordered by RID and then by Visit
  group_by(RID, Cohort) %>%
  summarise(
    NetVisits = n(), 
    FirstVisit = min(Visit), 
    LastVisit = max(Visit),
    AvgDurationBetweenVisits = mean(diff(Visit)), # Calculate average duration between visits
    .groups = 'drop' # Ensure the result is not grouped
  )

net_mean_repeat_visits = mean(summary_df$NetVisits, na.rm = TRUE)
net_SD_repeat_visits = sd(summary_df$NetVisits, na.rm = TRUE)
avg_duration_between_ScanVisits = mean(summary_df$AvgDurationBetweenVisits, na.rm = TRUE)
longest_participant_net = max(summary_df$LastVisit)
  
  
mean_repeat_visits_MCI_C = mean(summary_df$NetVisits[summary_df$Cohort=="MCI_C"], na.rm = TRUE)
SD_repeat_visits_MCI_C = sd(summary_df$NetVisits[summary_df$Cohort=="MCI_C"], na.rm = TRUE)
range_repeat_visits_MCI_C = range(summary_df$NetVisits[summary_df$Cohort=="MCI_C"], na.rm = TRUE)

mean_repeat_visits_MCI_NC = mean(summary_df$NetVisits[summary_df$Cohort=="MCI_NC"], na.rm = TRUE)
SD_repeat_visits_MCI_NC = sd(summary_df$NetVisits[summary_df$Cohort=="MCI_NC"], na.rm = TRUE)
range_repeat_visits_MCI_NC = range(summary_df$NetVisits[summary_df$Cohort=="MCI_NC"], na.rm = TRUE)

longest_participant_MCI_C = max(summary_df$LastVisit[summary_df$Cohort == "MCI_C"])
longest_participant_MCI_NC = max(summary_df$LastVisit[summary_df$Cohort == "MCI_NC"])

avg_duration_between_MCI_C_ScanVisits = mean(summary_df$AvgDurationBetweenVisits[summary_df$Cohort == "MCI_C"], na.rm = TRUE)

avg_duration_between_MCI_NC_ScanVisits = mean(summary_df$AvgDurationBetweenVisits[summary_df$Cohort == "MCI_NC"], na.rm = TRUE)
```

#Age of Onset
```{r}
#install.packages("dplyr")
library(dplyr)

# Assuming MCI is your original dataframe
new_dataframe <- MCI %>%
  filter(Dx_bl == "CN") %>%
  group_by(RID) %>%
  arrange(Visit, .by_group = TRUE) %>%
  slice(1) %>%
  select(RID, Age, Visit, Dx_bl, Diagnosis,Cohort)

```

`
#Number of subjects from each cohort
```{r}
library(dplyr)

# Assuming your dataframe is named MCI
count_rids_per_adni_protocol <- MCI %>%
  group_by(ADNI.Protocol) %>%
  summarise(UniqueRIDCount = n_distinct(RID))

# Display the result
print(count_rids_per_adni_protocol)
  
print(summary_df$RID[summary_df$NetVisits <=1])

net_participants = count_rids_per_adni_protocol

#Do any participants start in ADNI 3?
print(MCI$RID[MCI$Visit==0 && MCI$ADNI.Protocol =="ADNI3"])

```

#Install lmerTest
```{r}
#install.packages(lmerTest)
library(lmerTest)
install.packages("r2mlm")
library("r2mlm")

```


Making Cohort and Gender into factors (treated as categorical variables) + Relevelling:
```{r}
#Relevelling

#Cohort
MCI$Cohort = factor(MCI$Cohort)
cat("\nCurrent Levels:",levels(MCI$Cohort))
MCI$Cohort<- relevel(MCI$Cohort, "MCI_NC") #MCI_NC = 0, MCI_C = 1
cat("\nNew Levels:",levels(MCI$Cohort))

#Gender
MCI$Gender <- factor(MCI$Gender)
cat("\nGender Levels:",levels(MCI$Gender)) #Female = 0, Male = 1

```

#Retain only pre-conversion MCI visits

```{r}
MCI <- MCI[MCI$Diagnosis == "MCI",]
```

Standardizing with respect to Intracranial volume (instead of Whole Brain volume as previously done)

```{r}
#Standardizing Hippocampus and EC volume wrt ICV

MCI$Hippocampus_icv_absolute = MCI$Hippocampus/MCI$ICV
MCI$Entorhinal_icv_absolute = MCI$Entorhinal/MCI$ICV
MCI$Fusiform_icv_absolute = MCI$Fusiform/MCI$ICV
MCI$Ventricles_icv_absolute = MCI$Ventricles/MCI$ICV
MCI$Occipital_icv_absolute = MCI$Occipital/MCI$ICV
MCI$WholeBrain_icv_absolute = MCI$WholeBrain/MCI$ICV

#Storing SD and means of normalized Regional brain volumes
M_Hippocampus_icv = mean(MCI$Hippocampus_icv_absolute,na.rm = TRUE)
SD_Hippocampus_icv = sd(MCI$Hippocampus_icv_absolute,na.rm = TRUE)

M_Entorhinal_icv = mean(MCI$Entorhinal_icv_absolute,na.rm = TRUE)
S_Entorhinal_icv = sd(MCI$Entorhinal_icv_absolute,na.rm = TRUE)

M_Fusiform_icv = mean(MCI$Fusiform_icv_absolute,na.rm = TRUE)
SD_Fusiform_icv = sd(MCI$Fusiform_icv_absolute,na.rm = TRUE)

M_Occipital_icv = mean(MCI$Occipital_icv_absolute,na.rm = TRUE)
SD_Occipital_icv = sd(MCI$Occipital_icv_absolute,na.rm = TRUE)

M_Ventricles_icv = mean(MCI$Ventricles_icv_absolute,na.rm = TRUE)
SD_Ventricles_icv = sd(MCI$Ventricles_icv_absolute,na.rm = TRUE)

M_MWholeBrain_icv = mean(MCI$WholeBrain_icv_absolute,na.rm = TRUE)
SD_WholeBrain_icv = sd(MCI$WholeBrain_icv_absolute,na.rm = TRUE)

#Z-scoring variables
MCI$Hippocampus_icv = scale(MCI$Hippocampus_icv_absolute)
MCI$Entorhinal_icv = scale(MCI$Entorhinal_icv_absolute)
MCI$Fusiform_icv = scale(MCI$Fusiform_icv_absolute)
MCI$Ventricles_icv = scale(MCI$Ventricles_icv_absolute)
MCI$Occipital_icv = scale(MCI$Occipital_icv_absolute)
MCI$WholeBrain_icv = scale(MCI$WholeBrain_icv_absolute)

MCI$Age_sc = scale(MCI$Age)
M_Age = mean(MCI$Age,na.rm = TRUE)
SD_Age = sd(MCI$Age, na.rm = TRUE)
```


Fitting Growth Curves for Hippocampus volumes:

Creating an interaction term to look at stats:
```{r}
MCI$Hip_int = MCI$Age * as.numeric(MCI$Cohort)
MCI$Ent_int = MCI$Age * as.numeric(MCI$Cohort)

```

Scatterplot matrix for predictors:

```{r}
my_cols <- c("red", "blue")  
pairs(MCI[,c(90,89,6,92)], pch = 19,  cex = 0.5,
      col = my_cols[MCI$Cohort],
      lower.panel=NULL)

cor.test(as.numeric(MCI$Cohort),MCI$Hip_int)
```
#Since age is similarly distributed across cohorts, the binary predictor "cohort" and the interaction term "age * cohort" are bound to be highly correlated. Is this a multicollinearity problem?

# R treats levels of a binary factor as 1 and 2 instead of 0 and 1, as apparent from the interaction term (both MCI_C and MCI_NC are present). Is this a problem?


```{r}
#install necessary libraries
# install.packages('ggplot2')
# install.packages('GGally')

#load libraries
library(ggplot2)
library(GGally)

#create pairs plot
ggpairs(MCI[,c(90,89,6)])
```

Trial
```{r}
Hip_fixed <- lm(Hippocampus_icv ~ Age_sc + Cohort, data = MCI, na.action = na.omit)
summary(Hip_fixed)
```

Trial RE

```{r}
Hip_rnd_int <- lmer(Hippocampus_icv ~ Age + Cohort + (1 + Age|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = FALSE,na.action = na.exclude)
summary(Hip_rnd_int)
```



1. Only fixed effects of age and cohort:

```{r}
#Hippocampus growth curve models

Hip_fixed <- lm(Hippocampus_icv ~ Age_sc*Cohort, data = MCI, na.action = na.omit)
summary(Hip_fixed)
coef(Hip_fixed)
```
repeated measures anova:
```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)
```



2. Random intercepts for each subject, but fixed slope of age

```{r}
Hip_rnd_int <- lmer(Hippocampus_icv ~ Age_sc + Cohort + Age_sc : Cohort + (1|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = FALSE,na.action = na.exclude)
summary(Hip_rnd_int)
#coef(Hip_rnd_int)
```

3. Random Slopes + Random Intercept of age for each subject

```{r}
Hip_rnd_slope_int <- lmer(Hippocampus_icv ~ Age_sc + Cohort + ADNI.Protocol + Age_sc : Cohort + (1 + Age_sc|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = FALSE, na.action = na.exclude)
summary(Hip_rnd_slope_int)
#coef(Hip_rnd_slope_int)

```
*QUES Interpreting cohort fixed effect*

Correlations between random effects of intercept and slope
```{r}
 HRF =  ranef(Hip_rnd_slope_int)
plot(HRF$RID$`(Intercept)`,HRF$RID$Age, xlab = "Subject-Specific Deviations in Intercept", ylab = "Subject-Specific Deviations in Slope", main = "Correlation between Random Effects of \nIntercept and Slope - Hippocampus Trajectories")

```


Comparing the 3 models:

```{r}
r2mlm(Hip_rnd_slope_int)
#anova(Hip_rnd_int,Hip_rnd_slope_int,Hip_fixed)


```

------------------------------------------------------------------------------------------------------------------------
*QUES*

Interpretation:

AIC: fixed effects > random slopes only > random slopes + intercepts
Conclusion: Allowing individual intercepts and rates of change of Hippocampus volumes with age to deviate from the average is significantly improving model fit

BIC: fixed effects > random slopes only > random slopes + intercepts
Conclusion: Adding random effects for slope is not coming at the cost of parsimony of the model??
------------------------------------------------------------------------------------------------------------------------

#Stoing individual coefficients:
```{r}
#Coefficients
Hip_rand_eff <- ranef(Hip_rnd_slope_int)[["RID"]]
Hip_rand_eff <- cbind(RID=rownames(Hip_rand_eff),Hip_rand_eff)
#rownames(Hip_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

Hip_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(Hip_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (Hip_rand_eff[i,]$RID == MCI_C_ids[j])
    {Hip_rand_eff[i,]$Cohort = "MCI_C"}
  }}

```

#Computing final slopes for each individual (adding back fixed effects)

```{r}
#Adding back fixed effects to random effects to get final slope for each individual
Hip_rand_eff$Age_sc = Hip_rand_eff$Age_sc + Hip_rnd_slope_int@beta[2]  
Hip_rand_eff$`(Intercept)`= Hip_rand_eff$`(Intercept)`+ Hip_rnd_slope_int@beta[1]
Hip_rand_eff$Coh_slope = Hip_rnd_slope_int@beta[3]
Hip_rand_eff$Interaction = Hip_rnd_slope_int@beta[4]
```

---------------------------------------------------------------------------------------------------------------------------------------------
#Plotting Hippocampus rates of change:

```{r}
#fitted values
MCI$Hippocampus_fit = fitted(Hip_rnd_slope_int) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_Hippocampus_Coeff <- Hip_rand_eff[Hip_rand_eff$Cohort == "MCI_NC",]
MCI_C_Hippocampus_Coeff <- Hip_rand_eff[Hip_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (age) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Hippocampus_MCI_C_slope = mean(MCI_C_Hippocampus_Coeff$Age_sc, na.rm = TRUE)
Hippocampus_MCI_NC_slope = mean(MCI_NC_Hippocampus_Coeff$Age_sc, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Hippocampus_MCI_C_intercept = mean(MCI_C_Hippocampus_Coeff$`(Intercept)`, na.rm = TRUE)
Hippocampus_MCI_NC_intercept = mean(MCI_NC_Hippocampus_Coeff$`(Intercept)`, na.rm = TRUE)

#Add B2 (Cohort) to MCI_C intercept (as MCI_C is coded 1 and MCI_NC is coded 0)
Hippocampus_MCI_C_intercept = Hippocampus_MCI_C_intercept +  Hip_rnd_slope_int@beta[3]

#Add B3 (Interaction) for MCI_C slope (as MCI_C is coded 1 and MCI_NC is coded 0)
Hippocampus_MCI_C_slope = Hippocampus_MCI_C_slope + Hip_rnd_slope_int@beta[4]

ggplot(MCI, aes(x=Age_sc, y=Hippocampus_fit, group = RID, color = Cohort))+geom_line(size=1) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age_sc, y=Hippocampus_icv, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 1) + 
  geom_abline(slope = Hippocampus_MCI_C_slope, intercept = Hippocampus_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = Hippocampus_MCI_NC_slope, intercept = Hippocampus_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject (In Years)") + ylab("Hippocampal/Intracranial Volume (SD units)") + ggtitle("Changes in Hippocampus Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5))+ scale_x_continuous(breaks = c(((55-M_Age)/SD_Age),((60-M_Age)/SD_Age),((65-M_Age)/SD_Age),((70-M_Age)/SD_Age),((75-M_Age)/SD_Age),((80-M_Age)/SD_Age),((85-M_Age)/SD_Age),((90-M_Age)/SD_Age),((95-M_Age)/SD_Age)), labels = c(55,60,65,70,75,80,85,90,95 )) + theme_classic() + theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))




```


#Outlier is subject 2336, in one of their visits
MCI$Hippocampus[MCI$RID == 2336] (consistently decreasing)
MCI$ICV[MCI$RID == 2336] (randomly falls to around 2/3 of the average volume across visits in one visit!)

*REMOVE SUBJECT?*

 1192600 1189620 1196810 1194080     NaN  716133     NaN 1203970 1200870 1208220

---------------------------------------------------------------------------------------------------------------------------------------------

Fitting Growth Curves for Entorhinal Volume:

1. Only fixed effects model:

```{r}
#Ent_fixed <- model.matrix(Entorhinal_icv ~ Age*Cohort, data = MCI, na.action = na.omit)
Ent_fixed <- lm(Entorhinal_icv ~ Age_sc*Cohort, data = MCI, na.action = na.omit)

summary(Ent_fixed)
#coef(Ent_fixed)
```

*Commented out command indicates lm() automatically dummifies the binary variable - relevels it to 0 and 1, even though levels are naturally 1 and 2 in R*

2. Random intercepts + fixed slope model:

```{r}
Ent_rnd_int <- lmer(Entorhinal_icv ~ Age_sc + Cohort + Age_sc : Cohort + (1|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = FALSE,na.action = na.exclude)
summary(Ent_rnd_int)
#coef(Ent_rnd_int)
```

3. Random intercepts + Random slopes model:
```{r}
Ent_rnd_slope_int <- lmer(Entorhinal_icv ~ Age_sc + Cohort +ADNI.Protocol+ Age_sc : Cohort + (1 + Age_sc|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = FALSE,na.action = na.exclude)
summary(Ent_rnd_slope_int)
#coef(Ent_rnd_slope_int)

```
```{r}
anova(Ent_rnd_slope_int,Ent_rnd_int,Ent_fixed)

```

Correlations between random effects of intercept and slope
```{r}
ECRF =  ranef(Ent_rnd_slope_int)
plot(ECRF$RID$`(Intercept)`,ECRF$RID$Age, xlab = "Subject-Specific Deviations in Intercept", ylab = "Subject=Specific Deviations in Slope", main = "Correlation between Random Effects of \nIntercept and Slope - EC Trajectories")

```


#Dividing Coefficients up by Cohort:

#ranef returns only random effects (only individual deviations from fixed effects)

```{r}
#Coefficients
ent_rand_eff <- ranef(Ent_rnd_slope_int)[["RID"]]
ent_rand_eff <- cbind(RID=rownames(ent_rand_eff),ent_rand_eff)
#rownames(ent_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

ent_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(ent_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (ent_rand_eff[i,]$RID == MCI_C_ids[j])
    {ent_rand_eff[i,]$Cohort = "MCI_C"}
  }}


```

#Computing final slopes for each individual (adding back fixed effects)
```{r}

#Adding back fixed effects to random effects to get final slope for each individual
ent_rand_eff$Age_sc = ent_rand_eff$Age_sc + Ent_rnd_slope_int@beta[2]  
ent_rand_eff$`(Intercept)`= ent_rand_eff$`(Intercept)`+ Ent_rnd_slope_int@beta[1]
ent_rand_eff$Coh_slope = Ent_rnd_slope_int@beta[3]
ent_rand_eff$Interaction = Ent_rnd_slope_int@beta[4]

```

#Plotting Results:

```{r}

#fitted values
MCI$Entorhinal_fit = fitted(Ent_rnd_slope_int) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_Entorhinal_Coeff <- ent_rand_eff[ent_rand_eff$Cohort == "MCI_NC",]
MCI_C_Entorhinal_Coeff <- ent_rand_eff[ent_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (age) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Entorhinal_MCI_C_slope = mean(MCI_C_Entorhinal_Coeff$Age_sc, na.rm = TRUE)
Entorhinal_MCI_NC_slope = mean(MCI_NC_Entorhinal_Coeff$Age_sc, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Entorhinal_MCI_C_intercept = mean(MCI_C_Entorhinal_Coeff$`(Intercept)`, na.rm = TRUE)
Entorhinal_MCI_NC_intercept = mean(MCI_NC_Entorhinal_Coeff$`(Intercept)`, na.rm = TRUE)

#Add B2 (Cohort) to MCI_C intercept (as MCI_C is coded 1 and MCI_NC is coded 0)
Entorhinal_MCI_C_intercept = Entorhinal_MCI_C_intercept +  Ent_rnd_slope_int@beta[3]

#Add B3 (Interaction) for MCI_C slope (as MCI_C is coded 1 and MCI_NC is coded 0)
Entorhinal_MCI_C_slope = Entorhinal_MCI_C_slope + Ent_rnd_slope_int@beta[4]

ggplot(MCI, aes(x=Age_sc, y=Entorhinal_fit, group = RID, color = Cohort))+geom_line(size=1) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age_sc, y=Entorhinal_icv, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 1) + 
  geom_abline(slope = Entorhinal_MCI_C_slope, intercept = Entorhinal_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = Entorhinal_MCI_NC_slope, intercept = Entorhinal_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject (In Years)") + ylab("Entorhinal/Intracranial Volume (SD Units)") + ggtitle("Changes in Entorhinal Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5)) + scale_x_continuous(breaks = c(((55-M_Age)/SD_Age),((60-M_Age)/SD_Age),((65-M_Age)/SD_Age),((70-M_Age)/SD_Age),((75-M_Age)/SD_Age),((80-M_Age)/SD_Age),((85-M_Age)/SD_Age),((90-M_Age)/SD_Age),((95-M_Age)/SD_Age)), labels = c(55,60,65,70,75,80,85,90,95 )) +theme_classic() + theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))

```

T-tests for initial visits

```{r}
Hip =  data.frame(MCI$RID, MCI$Visit,MCI$Hippocampus_icv,MCI$Cohort)

Hip_1 = Hip[MCI$Visit==0,]
Hip_1_C = Hip_1[Hip_1$MCI.Cohort=="MCI_C",]
Hip_1_NC = Hip_1[Hip_1$MCI.Cohort=="MCI_NC",]
t.test(Hip_1_NC$MCI.Hippocampus_icv,Hip_1_C$MCI.Hippocampus_icv)

Ent =  data.frame(MCI$RID, MCI$Visit,MCI$Entorhinal_icv,MCI$Cohort)

Ent_1 = Ent[MCI$Visit==0,]
Ent_1_C = Ent_1[Ent_1$MCI.Cohort=="MCI_C",]
Ent_1_NC = Ent_1[Ent_1$MCI.Cohort=="MCI_NC",]
t.test(Ent_1_NC$MCI.Entorhinal_icv,Ent_1_C$MCI.Entorhinal_icv)
```

#FUSIFORM Gyrus Volume MODEL

```{r}
#Ent_fixed <- model.matrix(Entorhinal_icv ~ Age*Cohort, data = MCI, na.action = na.omit)
Fus_fixed <- lm(Fusiform_icv ~ Age_sc*Cohort, data = MCI, na.action = na.omit)

summary(Fus_fixed)
#coef(Ent_fixed)
```

```{r}
# Random Slopes + Random Intercept of age for each subject

Fusif_rnd_slope_int <- lmer(Fusiform_icv ~ Age_sc + Cohort +ADNI.Protocol+ Age_sc : Cohort + (1 + Age_sc|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = FALSE, na.action = na.exclude)
summary(Fusif_rnd_slope_int)

```


```{r}

#Stoing individual coefficients:

#Coefficients
Fusif_rand_eff <- ranef(Fusif_rnd_slope_int)[["RID"]]
Fusif_rand_eff <- cbind(RID=rownames(Fusif_rand_eff),Fusif_rand_eff)
#rownames(Fusif_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

Fusif_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(Fusif_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (Fusif_rand_eff[i,]$RID == MCI_C_ids[j])
    {Fusif_rand_eff[i,]$Cohort = "MCI_C"}
  }}

#Computing final slopes for each individual (adding back fixed effects)


#Adding back fixed effects to random effects to get final slope for each individual
Fusif_rand_eff$Age_sc = Fusif_rand_eff$Age_sc + Fusif_rnd_slope_int@beta[2]  
Fusif_rand_eff$`(Intercept)`= Fusif_rand_eff$`(Intercept)`+ Fusif_rnd_slope_int@beta[1]
Fusif_rand_eff$Coh_slope = Fusif_rnd_slope_int@beta[3]
Fusif_rand_eff$Interaction = Fusif_rnd_slope_int@beta[4]


#Plotting Fusiform rates of change
#fitted values
MCI$Fusiform_fit = fitted(Fusif_rnd_slope_int) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_Fusiform_Coeff <- Fusif_rand_eff[Fusif_rand_eff$Cohort == "MCI_NC",]
MCI_C_Fusiform_Coeff <- Fusif_rand_eff[Fusif_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (age) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Fusiform_MCI_C_slope = mean(MCI_C_Fusiform_Coeff$Age_sc, na.rm = TRUE)
Fusiform_MCI_NC_slope = mean(MCI_NC_Fusiform_Coeff$Age_sc, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Fusiform_MCI_C_intercept = mean(MCI_C_Fusiform_Coeff$`(Intercept)`, na.rm = TRUE)
Fusiform_MCI_NC_intercept = mean(MCI_NC_Fusiform_Coeff$`(Intercept)`, na.rm = TRUE)

#Add B2 (Cohort) to MCI_C intercept (as MCI_C is coded 1 and MCI_NC is coded 0)
Fusiform_MCI_C_intercept = Fusiform_MCI_C_intercept +  Fusif_rnd_slope_int@beta[3]

#Add B3 (Interaction) for MCI_C slope (as MCI_C is coded 1 and MCI_NC is coded 0)
Fusiform_MCI_C_slope = Fusiform_MCI_C_slope + Fusif_rnd_slope_int@beta[4]

```
 
#Fusiform Plot - WEAKER age cohort interaction than entorhinal? Still significant
```{r}
ggplot(MCI, aes(x=Age_sc, y=Fusiform_fit, group = RID, color = Cohort))+geom_line(size=1) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age_sc, y=Fusiform_icv, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 1) + 
  geom_abline(slope = Fusiform_MCI_C_slope, intercept = Fusiform_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = Fusiform_MCI_NC_slope, intercept = Fusiform_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject") + ylab("Fusiform Gyrus/Intracranial Volume (SD units)") + ggtitle("Changes in Fusiform Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5)) + scale_x_continuous(breaks = c(((55-M_Age)/SD_Age),((60-M_Age)/SD_Age),((65-M_Age)/SD_Age),((70-M_Age)/SD_Age),((75-M_Age)/SD_Age),((80-M_Age)/SD_Age),((85-M_Age)/SD_Age),((90-M_Age)/SD_Age),((95-M_Age)/SD_Age)), labels = c(55,60,65,70,75,80,85,90,95 )) + theme_classic() + theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))

```

#VENTRICLES MODEL

```{r}
# Random Slopes + Random Intercept of age for each subject

Ven_rnd_slope_int <- lmer(Ventricles_icv ~ Age_sc + Cohort + ADNI.Protocol + Age_sc : Cohort + (1 + Age_sc|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = FALSE, na.action = na.exclude)
summary(Ven_rnd_slope_int)

```

```{r}
#Stoing individual coefficients:

#Coefficients
Ven_rand_eff <- ranef(Ven_rnd_slope_int)[["RID"]]
Ven_rand_eff <- cbind(RID=rownames(Ven_rand_eff),Ven_rand_eff)
#rownames(Ven_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

Ven_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(Ven_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (Ven_rand_eff[i,]$RID == MCI_C_ids[j])
    {Ven_rand_eff[i,]$Cohort = "MCI_C"}
  }}

#Computing final slopes for each individual (adding back fixed effects)


#Adding back fixed effects to random effects to get final slope for each individual
Ven_rand_eff$Age_sc = Ven_rand_eff$Age_sc + Ven_rnd_slope_int@beta[2]  
Ven_rand_eff$`(Intercept)`= Ven_rand_eff$`(Intercept)`+ Ven_rnd_slope_int@beta[1]
Ven_rand_eff$Coh_slope = Ven_rnd_slope_int@beta[3]
Ven_rand_eff$Interaction = Ven_rnd_slope_int@beta[4]


#Plotting Ventricles rates of change
#fitted values
MCI$Ventricles_fit = fitted(Ven_rnd_slope_int) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_Ventricles_Coeff <- Ven_rand_eff[Ven_rand_eff$Cohort == "MCI_NC",]
MCI_C_Ventricles_Coeff <- Ven_rand_eff[Ven_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (age) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Ventricles_MCI_C_slope = mean(MCI_C_Ventricles_Coeff$Age_sc, na.rm = TRUE)
Ventricles_MCI_NC_slope = mean(MCI_NC_Ventricles_Coeff$Age_sc, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Ventricles_MCI_C_intercept = mean(MCI_C_Ventricles_Coeff$`(Intercept)`, na.rm = TRUE)
Ventricles_MCI_NC_intercept = mean(MCI_NC_Ventricles_Coeff$`(Intercept)`, na.rm = TRUE)

#Add B2 (Cohort) to MCI_C intercept (as MCI_C is coded 1 and MCI_NC is coded 0)
Ventricles_MCI_C_intercept = Ventricles_MCI_C_intercept +  Ven_rnd_slope_int@beta[3]

#Add B3 (Interaction) for MCI_C slope (as MCI_C is coded 1 and MCI_NC is coded 0)
Ventricles_MCI_C_slope = Ventricles_MCI_C_slope + Ven_rnd_slope_int@beta[4]


```

#VENTRICLES PLOT - Pretty Strong age-cohort interaction 

```{r}
ggplot(MCI, aes(x=Age_sc, y=Ventricles_fit, group = RID, color = Cohort))+geom_line(size=1) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age_sc, y=Ventricles_icv, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 1) + 
  geom_abline(slope = Ventricles_MCI_C_slope, intercept = Ventricles_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = Ventricles_MCI_NC_slope, intercept = Ventricles_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject") + ylab("Ventricles/Intracranial Volume (SD units)") + ggtitle("Changes in Ventricle Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5)) +scale_x_continuous(breaks = c(((55-M_Age)/SD_Age),((60-M_Age)/SD_Age),((65-M_Age)/SD_Age),((70-M_Age)/SD_Age),((75-M_Age)/SD_Age),((80-M_Age)/SD_Age),((85-M_Age)/SD_Age),((90-M_Age)/SD_Age),((95-M_Age)/SD_Age)), labels = c(55,60,65,70,75,80,85,90,95 )) + theme_classic() + theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))
```

```{r}
# Random Slopes + Random Intercept of age for each subject

Occip_rnd_slope_int <- lmer(Occipital_icv ~ Age_sc + Cohort + ADNI.Protocol + Age_sc : Cohort + (1 + Age_sc|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = FALSE, na.action = na.exclude)
summary(Occip_rnd_slope_int)

```


```{r}

#Storing individual coefficients:

#Coefficients
Occip_rand_eff <- ranef(Occip_rnd_slope_int)[["RID"]]
Occip_rand_eff <- cbind(RID=rownames(Occip_rand_eff),Occip_rand_eff)
#rownames(Occip_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

Occip_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(Occip_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (Occip_rand_eff[i,]$RID == MCI_C_ids[j])
    {Occip_rand_eff[i,]$Cohort = "MCI_C"}
  }}

#Computing final slopes for each individual (adding back fixed effects)


#Adding back fixed effects to random effects to get final slope for each individual
Occip_rand_eff$Age_sc = Occip_rand_eff$Age_sc + Occip_rnd_slope_int@beta[2]  
Occip_rand_eff$`(Intercept)`= Occip_rand_eff$`(Intercept)`+ Occip_rnd_slope_int@beta[1]
Occip_rand_eff$Coh_slope = Occip_rnd_slope_int@beta[3]
Occip_rand_eff$Interaction = Occip_rnd_slope_int@beta[4]


#Plotting Occipital rates of change
#fitted values
MCI$Occipital_fit = fitted(Occip_rnd_slope_int) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_Occipital_Coeff <- Occip_rand_eff[Occip_rand_eff$Cohort == "MCI_NC",]
MCI_C_Occipital_Coeff <- Occip_rand_eff[Occip_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (age) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Occipital_MCI_C_slope = mean(MCI_C_Occipital_Coeff$Age_sc, na.rm = TRUE)
Occipital_MCI_NC_slope = mean(MCI_NC_Occipital_Coeff$Age_sc, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Occipital_MCI_C_intercept = mean(MCI_C_Occipital_Coeff$`(Intercept)`, na.rm = TRUE)
Occipital_MCI_NC_intercept = mean(MCI_NC_Occipital_Coeff$`(Intercept)`, na.rm = TRUE)

#Add B2 (Cohort) to MCI_C intercept (as MCI_C is coded 1 and MCI_NC is coded 0)
Occipital_MCI_C_intercept = Occipital_MCI_C_intercept +  Occip_rnd_slope_int@beta[3]

#Add B3 (Interaction) for MCI_C slope (as MCI_C is coded 1 and MCI_NC is coded 0)
Occipital_MCI_C_slope = Occipital_MCI_C_slope + Occip_rnd_slope_int@beta[4]

```

#Occipital Plot - WEAKER age cohort interaction than entorhinal? Still significant
```{r}
ggplot(MCI, aes(x=Age_sc, y=Occipital_fit, group = RID, color = Cohort))+geom_line(size=1) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age_sc, y=Occipital_icv, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 1) + 
  geom_abline(slope = Occipital_MCI_C_slope, intercept = Occipital_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = Occipital_MCI_NC_slope, intercept = Occipital_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject") + ylab("Occipital/Intracranial Volume (SD units)") + ggtitle("Changes in Occipital Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5)) +scale_x_continuous(breaks = c(((55-M_Age)/SD_Age),((60-M_Age)/SD_Age),((65-M_Age)/SD_Age),((70-M_Age)/SD_Age),((75-M_Age)/SD_Age),((80-M_Age)/SD_Age),((85-M_Age)/SD_Age),((90-M_Age)/SD_Age),((95-M_Age)/SD_Age)), labels = c(55,60,65,70,75,80,85,90,95 )) +theme_classic() + theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))
```


```{r}
# Random Slopes + Random Intercept of age for each subject

WB_rnd_slope_int <- lmer(WholeBrain_icv ~ Age_sc + Cohort + Age_sc : Cohort + (1 + Age_sc|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = FALSE, na.action = na.exclude)
summary(WB_rnd_slope_int)

```

```{r}

#Storing individual coefficients:

#Coefficients
WB_rand_eff <- ranef(WB_rnd_slope_int)[["RID"]]
WB_rand_eff <- cbind(RID=rownames(WB_rand_eff),WB_rand_eff)
#rownames(WB_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

WB_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(WB_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (WB_rand_eff[i,]$RID == MCI_C_ids[j])
    {WB_rand_eff[i,]$Cohort = "MCI_C"}
  }}

#Computing final slopes for each individual (adding back fixed effects)

#Adding back fixed effects to random effects to get final slope for each individual
WB_rand_eff$Age_sc = WB_rand_eff$Age_sc + WB_rnd_slope_int@beta[2]  
WB_rand_eff$`(Intercept)`= WB_rand_eff$`(Intercept)`+ WB_rnd_slope_int@beta[1]
WB_rand_eff$Coh_slope = WB_rnd_slope_int@beta[3]
WB_rand_eff$Interaction = WB_rnd_slope_int@beta[4]


#Plotting WholeBrain rates of change
#fitted values
MCI$WholeBrain_fit = fitted(WB_rnd_slope_int) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_WholeBrain_Coeff <- WB_rand_eff[WB_rand_eff$Cohort == "MCI_NC",]
MCI_C_WholeBrain_Coeff <- WB_rand_eff[WB_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (age) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
WholeBrain_MCI_C_slope = mean(MCI_C_WholeBrain_Coeff$Age_sc, na.rm = TRUE)
WholeBrain_MCI_NC_slope = mean(MCI_NC_WholeBrain_Coeff$Age_sc, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
WholeBrain_MCI_C_intercept = mean(MCI_C_WholeBrain_Coeff$`(Intercept)`, na.rm = TRUE)
WholeBrain_MCI_NC_intercept = mean(MCI_NC_WholeBrain_Coeff$`(Intercept)`, na.rm = TRUE)

#Add B2 (Cohort) to MCI_C intercept (as MCI_C is coded 1 and MCI_NC is coded 0)
WholeBrain_MCI_C_intercept = WholeBrain_MCI_C_intercept +  WB_rnd_slope_int@beta[3]

#Add B3 (Interaction) for MCI_C slope (as MCI_C is coded 1 and MCI_NC is coded 0)
WholeBrain_MCI_C_slope = WholeBrain_MCI_C_slope + WB_rnd_slope_int@beta[4]

```

#WholeBrain Plot - WEAKER age cohort interaction than entorhinal? Still significant
```{r}
ggplot(MCI, aes(x=Age_sc, y=WholeBrain_fit, group = RID, color = Cohort))+geom_line(size=1) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age_sc, y=WholeBrain_icv, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 1) + 
  geom_abline(slope = WholeBrain_MCI_C_slope, intercept = WholeBrain_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = WholeBrain_MCI_NC_slope, intercept = WholeBrain_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject") + ylab("WholeBrain/Intracranial Volume (SD units)") + ggtitle("Changes in WholeBrain Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5)) + scale_x_continuous(breaks = c(((55-M_Age)/SD_Age),((60-M_Age)/SD_Age),((65-M_Age)/SD_Age),((70-M_Age)/SD_Age),((75-M_Age)/SD_Age),((80-M_Age)/SD_Age),((85-M_Age)/SD_Age),((90-M_Age)/SD_Age),((95-M_Age)/SD_Age)), labels = c(55,60,65,70,75,80,85,90,95 )) + theme_classic() + theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))

```

ADDED CHECKS - Nov 23
Hippocampus Unstandradized Rates
```{r}
# Random Slopes + Random Intercept of age for each subject

Occip_rnd_slope_int <- lmer(Occipital ~ Age + Cohort + ADNI.Protocol + Age : Cohort + (1 + Age|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = FALSE, na.action = na.exclude)
summary(Occip_rnd_slope_int)

```


```{r}

#Storing individual coefficients:

#Coefficients
Occip_rand_eff <- ranef(Occip_rnd_slope_int)[["RID"]]
Occip_rand_eff <- cbind(RID=rownames(Occip_rand_eff),Occip_rand_eff)
#rownames(Occip_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

Occip_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(Occip_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (Occip_rand_eff[i,]$RID == MCI_C_ids[j])
    {Occip_rand_eff[i,]$Cohort = "MCI_C"}
  }}

#Computing final slopes for each individual (adding back fixed effects)


#Adding back fixed effects to random effects to get final slope for each individual
Occip_rand_eff$Age = Occip_rand_eff$Age + Occip_rnd_slope_int@beta[2]  
Occip_rand_eff$`(Intercept)`= Occip_rand_eff$`(Intercept)`+ Occip_rnd_slope_int@beta[1]
Occip_rand_eff$Coh_slope = Occip_rnd_slope_int@beta[3]
Occip_rand_eff$Interaction = Occip_rnd_slope_int@beta[4]


#Plotting Occipital rates of change
#fitted values
MCI$Occipital_fit = fitted(Occip_rnd_slope_int) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_Un_Occipital_Coeff <- Occip_rand_eff[Occip_rand_eff$Cohort == "MCI_NC",]
MCI_C_Un_Occipital_Coeff <- Occip_rand_eff[Occip_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (age) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Un_Occipital_MCI_C_slope = mean(MCI_C_Un_Occipital_Coeff$Age, na.rm = TRUE)
Un_Occipital_MCI_NC_slope = mean(MCI_NC_Un_Occipital_Coeff$Age, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Un_Occipital_MCI_C_intercept = mean(MCI_C_Un_Occipital_Coeff$`(Intercept)`, na.rm = TRUE)
Un_Occipital_MCI_NC_intercept = mean(MCI_NC_Un_Occipital_Coeff$`(Intercept)`, na.rm = TRUE)

#Add B2 (Cohort) to MCI_C intercept (as MCI_C is coded 1 and MCI_NC is coded 0)
Un_Occipital_MCI_C_intercept = Un_Occipital_MCI_C_intercept +  Occip_rnd_slope_int@beta[3]

#Add B3 (Interaction) for MCI_C slope (as MCI_C is coded 1 and MCI_NC is coded 0)
Un_Occipital_MCI_C_slope = Un_Occipital_MCI_C_slope + Occip_rnd_slope_int@beta[4]

```

#Occipital Plot - WEAKER age cohort interaction than entorhinal? Still significant
```{r}
ggplot(MCI, aes(x=Age, y=Occipital_fit, group = RID, color = Cohort))+geom_line(size=1.5) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age, y=Occipital, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 2) + 
  geom_abline(slope = Un_Occipital_MCI_C_slope, intercept = Un_Occipital_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = Un_Occipital_MCI_NC_slope, intercept = Un_Occipital_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject") + ylab("Occipital Volume (mm3)") + ggtitle("Changes in Occipital Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5)) +scale_x_continuous(breaks = c(((55-M_Age)/SD_Age),((60-M_Age)/SD_Age),((65-M_Age)/SD_Age),((70-M_Age)/SD_Age),((75-M_Age)/SD_Age),((80-M_Age)/SD_Age),((85-M_Age)/SD_Age),((90-M_Age)/SD_Age),((95-M_Age)/SD_Age)), labels = c(55,60,65,70,75,80,85,90,95 )) +theme_classic() + theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))
```


ICV Rates
```{r}
# Random Slopes + Random Intercept of age for each subject

ICV_rnd_slope_int <- lmer(ICV ~ Age + Cohort + Age : Cohort + (1 + Age|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = FALSE, na.action = na.exclude)
summary(ICV_rnd_slope_int)

```


```{r}

#Storing individual coefficients:

#Coefficients
ICV_rand_eff <- ranef(ICV_rnd_slope_int)[["RID"]]
ICV_rand_eff <- cbind(RID=rownames(ICV_rand_eff),ICV_rand_eff)
#rownames(ICV_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

ICV_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(ICV_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (ICV_rand_eff[i,]$RID == MCI_C_ids[j])
    {ICV_rand_eff[i,]$Cohort = "MCI_C"}
  }}

#Computing final slopes for each individual (adding back fixed effects)


#Adding back fixed effects to random effects to get final slope for each individual
ICV_rand_eff$Age = ICV_rand_eff$Age + ICV_rnd_slope_int@beta[2]  
ICV_rand_eff$`(Intercept)`= ICV_rand_eff$`(Intercept)`+ ICV_rnd_slope_int@beta[1]
ICV_rand_eff$Coh_slope = ICV_rnd_slope_int@beta[3]
ICV_rand_eff$Interaction = ICV_rnd_slope_int@beta[4]


#Plotting ICV rates of change
#fitted values
MCI$ICV_fit = fitted(ICV_rnd_slope_int) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_Un_ICV_Coeff <- ICV_rand_eff[ICV_rand_eff$Cohort == "MCI_NC",]
MCI_C_Un_ICV_Coeff <- ICV_rand_eff[ICV_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (age) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Un_ICV_MCI_C_slope = mean(MCI_C_Un_ICV_Coeff$Age, na.rm = TRUE)
Un_ICV_MCI_NC_slope = mean(MCI_NC_Un_ICV_Coeff$Age, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Un_ICV_MCI_C_intercept = mean(MCI_C_Un_ICV_Coeff$`(Intercept)`, na.rm = TRUE)
Un_ICV_MCI_NC_intercept = mean(MCI_NC_Un_ICV_Coeff$`(Intercept)`, na.rm = TRUE)

#Add B2 (Cohort) to MCI_C intercept (as MCI_C is coded 1 and MCI_NC is coded 0)
Un_ICV_MCI_C_intercept = Un_ICV_MCI_C_intercept +  ICV_rnd_slope_int@beta[3]

#Add B3 (Interaction) for MCI_C slope (as MCI_C is coded 1 and MCI_NC is coded 0)
Un_ICV_MCI_C_slope = Un_ICV_MCI_C_slope + ICV_rnd_slope_int@beta[4]

```

#ICV Plot - WEAKER age cohort interaction than entorhinal? Still significant
```{r}
ggplot(MCI, aes(x=Age, y=ICV_fit, group = RID, color = Cohort))+geom_line(size=1.5) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age, y=ICV, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 2) + 
  geom_abline(slope = Un_ICV_MCI_C_slope, intercept = Un_ICV_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = Un_ICV_MCI_NC_slope, intercept = Un_ICV_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject") + ylab("ICV Volume (mm3)") + ggtitle("Changes in ICV Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5)) +scale_x_continuous(breaks = c(((55-M_Age)/SD_Age),((60-M_Age)/SD_Age),((65-M_Age)/SD_Age),((70-M_Age)/SD_Age),((75-M_Age)/SD_Age),((80-M_Age)/SD_Age),((85-M_Age)/SD_Age),((90-M_Age)/SD_Age),((95-M_Age)/SD_Age)), labels = c(55,60,65,70,75,80,85,90,95 )) +theme_classic() + theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))
```










