---
title: "GC for only Pre-Conversion Visits"
output: html_notebook
---

---
title: "ADNI Analysis 1 : Fitting Growth Curves"
---
Redaing in Data:
```{r}
#Reading in Data
#MCI<- data.frame(read.table("/Users/shreyarajagopal/Documents/Research/ADNI_4/Subject_Selection/MCI_Only_MCI_AD_Visits_5_March22.csv",header = T, sep = ","))

#Dataset with occipital volume
MCI <-data.frame(read.table("/Users/shreyarajagopal/Documents/Research/ADNI_4/Subject_Selection/MCI_Jun22.csv",header = T, sep = ","))
#Number of subjects
cat("MCI_C Number of Subjects: ",length(unique(MCI$RID[MCI$Cohort == "MCI_C"]))) 
cat("\nMCI_NC Number of Subjects: ",length(unique(MCI$RID[MCI$Cohort == "MCI_NC"]))) 
```

```{r}
MCI = MCI[MCI$RID!=2336,] #Removing subjects due to issues with ICV measurements
#MCI = MCI[MCI$RID!=6632,] #Removing subject due to issues with MMSE=22 and CDRSB = 5
#MCI = MCI[MCI$RID!=6341,] #Removing subject due to issues with MMSE = 19
```

#Install lmerTest
```{r}
#install.packages(lmerTest)
library(lmerTest)
```

#Retain only pre-conversion MCI visits

```{r}
MCI <- MCI[MCI$Diagnosis == "MCI",]
```

#Making Cohort and Gender into factors (treated as categorical variables) + Relevelling:
```{r}
#Relevelling

#Cohort
MCI$Cohort = factor(MCI$Cohort)
cat("\nCurrent Levels:",levels(MCI$Cohort))
MCI$Cohort<- relevel(MCI$Cohort, "MCI_NC") #MCI_NC = 0, MCI_C = 1
cat("\nNew Levels:",levels(MCI$Cohort))

#Gender
MCI$Gender <- factor(MCI$Gender)
cat("\nGender Levels:",levels(MCI$Gender)) #Female = 0, Male = 1

```


Standardizing with respect to Intracranial volume (instead of Whole Brain volume as previously done)

```{r}
#Standardizing Hippocampus and EC volume wrt ICV

MCI$Hippocampus_icv = MCI$Hippocampus/MCI$ICV
MCI$Entorhinal_icv = MCI$Entorhinal/MCI$ICV
MCI$Fusiform_icv = MCI$Fusiform/MCI$ICV
MCI$Ventricles_icv = MCI$Ventricles/MCI$ICV
MCI$Occipital_icv = MCI$Occipital/MCI$ICV
MCI$WholeBrain_icv = MCI$WholeBrain/MCI$ICV

#Z-scoring variables
MCI$Hippocampus_icv = scale(MCI$Hippocampus_icv)
MCI$Entorhinal_icv = scale(MCI$Entorhinal_icv)
MCI$Fusiform_icv = scale(MCI$Fusiform_icv)
MCI$Ventricles_icv = scale(MCI$Ventricles_icv)
MCI$Occipital_icv = scale(MCI$Occipital_icv)
MCI$WholeBrain_icv = scale(MCI$WholeBrain_icv)
```


Fitting Growth Curves for Hippocampus volumes:

Creating an interaction term to look at stats:
```{r}
MCI$Hip_int = MCI$Age * as.numeric(MCI$Cohort)
MCI$Ent_int = MCI$Age * as.numeric(MCI$Cohort)

```

Scatterplot matrix for predictors:

```{r}
my_cols <- c("red", "blue")  
pairs(MCI[,c(90,89,6,92)], pch = 19,  cex = 0.5,
      col = my_cols[MCI$Cohort],
      lower.panel=NULL)

cor.test(as.numeric(MCI$Cohort),MCI$Hip_int)
```
#Since age is similarly distributed across cohorts, the binary predictor "cohort" and the interaction term "age * cohort" are bound to be highly correlated. Is this a multicollinearity problem?

# R treats levels of a binary factor as 1 and 2 instead of 0 and 1, as apparent from the interaction term (both MCI_C and MCI_NC are present). Is this a problem?


```{r}
#install necessary libraries
# install.packages('ggplot2')
# install.packages('GGally')

#load libraries
library(ggplot2)
library(GGally)

#create pairs plot
ggpairs(MCI[,c(90,89,6)])
```

Trial
```{r}
Hip_fixed <- lm(Hippocampus_icv ~ Age + Cohort, data = MCI, na.action = na.omit)
summary(Hip_fixed)
```

Trial RE

```{r}
Hip_rnd_int <- lmer(Hippocampus_icv ~ Age + Cohort + (1 + Age|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = FALSE,na.action = na.exclude)
summary(Hip_rnd_int)
```



1. Only fixed effects of age and cohort:

```{r}
#Hippocampus growth curve models

Hip_fixed <- lm(Hippocampus_icv ~ Age*Cohort, data = MCI, na.action = na.omit)
summary(Hip_fixed)
coef(Hip_fixed)
```

2. Random intercepts for each subject, but fixed slope of age

```{r}
Hip_rnd_int <- lmer(Hippocampus_icv ~ Age + Cohort + Age : Cohort + (1|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = FALSE,na.action = na.exclude)
summary(Hip_rnd_int)
#coef(Hip_rnd_int)
```

3. Random Slopes + Random Intercept of age for each subject

```{r}
Hip_rnd_slope_int <- lmer(Hippocampus_icv ~ Age + Cohort + Age : Cohort + (1 + Age|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = FALSE, na.action = na.exclude)
summary(Hip_rnd_slope_int)
#coef(Hip_rnd_slope_int)

```
*QUES Interpreting cohort fixed effect*

Correlations between random effects of intercept and slope
```{r}
 HRF =  ranef(Hip_rnd_slope_int)
plot(HRF$RID$`(Intercept)`,HRF$RID$Age, xlab = "Subject-Specific Deviations in Intercept", ylab = "Subject-Specific Deviations in Slope", main = "Correlation between Random Effects of \nIntercept and Slope - Hippocampus Trajectories")

```


Comparing the 3 models:

```{r}
anova(Hip_rnd_int,Hip_rnd_slope_int,Hip_fixed)

```

------------------------------------------------------------------------------------------------------------------------
*QUES*

Interpretation:

AIC: fixed effects > random slopes only > random slopes + intercepts
Conclusion: Allowing individual intercepts and rates of change of Hippocampus volumes with age to deviate from the average is significantly improving model fit

BIC: fixed effects > random slopes only > random slopes + intercepts
Conclusion: Adding random effects for slope is not coming at the cost of parsimony of the model??
------------------------------------------------------------------------------------------------------------------------

#Stoing individual coefficients:
```{r}
#Coefficients
Hip_rand_eff <- ranef(Hip_rnd_slope_int)[["RID"]]
Hip_rand_eff <- cbind(RID=rownames(Hip_rand_eff),Hip_rand_eff)
#rownames(Hip_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

Hip_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(Hip_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (Hip_rand_eff[i,]$RID == MCI_C_ids[j])
    {Hip_rand_eff[i,]$Cohort = "MCI_C"}
  }}

```

#Computing final slopes for each individual (adding back fixed effects)

```{r}
#Adding back fixed effects to random effects to get final slope for each individual
Hip_rand_eff$Age = Hip_rand_eff$Age + Hip_rnd_slope_int@beta[2]  
Hip_rand_eff$`(Intercept)`= Hip_rand_eff$`(Intercept)`+ Hip_rnd_slope_int@beta[1]
Hip_rand_eff$Coh_slope = Hip_rnd_slope_int@beta[3]
Hip_rand_eff$Interaction = Hip_rnd_slope_int@beta[4]
```

---------------------------------------------------------------------------------------------------------------------------------------------
#Plotting Hippocampus rates of change:

```{r}
#fitted values
MCI$Hippocampus_fit = fitted(Hip_rnd_slope_int) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_Hippocampus_Coeff <- Hip_rand_eff[Hip_rand_eff$Cohort == "MCI_NC",]
MCI_C_Hippocampus_Coeff <- Hip_rand_eff[Hip_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (age) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Hippocampus_MCI_C_slope = mean(MCI_C_Hippocampus_Coeff$Age, na.rm = TRUE)
Hippocampus_MCI_NC_slope = mean(MCI_NC_Hippocampus_Coeff$Age, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Hippocampus_MCI_C_intercept = mean(MCI_C_Hippocampus_Coeff$`(Intercept)`, na.rm = TRUE)
Hippocampus_MCI_NC_intercept = mean(MCI_NC_Hippocampus_Coeff$`(Intercept)`, na.rm = TRUE)

#Add B2 (Cohort) to MCI_C intercept (as MCI_C is coded 1 and MCI_NC is coded 0)
Hippocampus_MCI_C_intercept = Hippocampus_MCI_C_intercept +  Hip_rnd_slope_int@beta[3]

#Add B3 (Interaction) for MCI_C slope (as MCI_C is coded 1 and MCI_NC is coded 0)
Hippocampus_MCI_C_slope = Hippocampus_MCI_C_slope + Hip_rnd_slope_int@beta[4]

ggplot(MCI, aes(x=Age, y=Hippocampus_fit, group = RID, color = Cohort))+geom_line(size=1.5) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age, y=Hippocampus_icv, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 2) + 
  geom_abline(slope = Hippocampus_MCI_C_slope, intercept = Hippocampus_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = Hippocampus_MCI_NC_slope, intercept = Hippocampus_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject") + ylab("Standardized Hippocampus Volume (SD units)") + ggtitle("Changes in Hippocampus Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5))  

```


#Outlier is subject 2336, in one of their visits
MCI$Hippocampus[MCI$RID == 2336] (consistently decreasing)
MCI$ICV[MCI$RID == 2336] (randomly falls to around 2/3 of the average volume across visits in one visit!)

*REMOVE SUBJECT?*

 1192600 1189620 1196810 1194080     NaN  716133     NaN 1203970 1200870 1208220

---------------------------------------------------------------------------------------------------------------------------------------------

Fitting Growth Curves for Entorhinal Volume:

1. Only fixed effects model:

```{r}
#Ent_fixed <- model.matrix(Entorhinal_icv ~ Age*Cohort, data = MCI, na.action = na.omit)
Ent_fixed <- lm(Entorhinal_icv ~ Age*Cohort, data = MCI, na.action = na.omit)

summary(Ent_fixed)
#coef(Ent_fixed)
```

*Commented out command indicates lm() automatically dummifies the binary variable - relevels it to 0 and 1, even though levels are naturally 1 and 2 in R*

2. Random intercepts + fixed slope model:

```{r}
Ent_rnd_int <- lmer(Entorhinal_icv ~ Age + Cohort + Age : Cohort + (1|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = FALSE,na.action = na.exclude)
summary(Ent_rnd_int)
#coef(Ent_rnd_int)
```

3. Random intercepts + Random slopes model:
```{r}
Ent_rnd_slope_int <- lmer(Entorhinal_icv ~ Age + Cohort + Age : Cohort + (1 + Age|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = FALSE,na.action = na.exclude)
summary(Ent_rnd_slope_int)
#coef(Ent_rnd_slope_int)

```
```{r}
anova(Ent_rnd_slope_int,Ent_rnd_int,Ent_fixed)

```

Correlations between random effects of intercept and slope
```{r}
ECRF =  ranef(Ent_rnd_slope_int)
plot(ECRF$RID$`(Intercept)`,ECRF$RID$Age, xlab = "Subject-Specific Deviations in Intercept", ylab = "Subject=Specific Deviations in Slope", main = "Correlation between Random Effects of \nIntercept and Slope - EC Trajectories")

```


#Dividing Coefficients up by Cohort:

#ranef returns only random effects (only individual deviations from fixed effects)

```{r}
#Coefficients
ent_rand_eff <- ranef(Ent_rnd_slope_int)[["RID"]]
ent_rand_eff <- cbind(RID=rownames(ent_rand_eff),ent_rand_eff)
#rownames(ent_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

ent_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(ent_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (ent_rand_eff[i,]$RID == MCI_C_ids[j])
    {ent_rand_eff[i,]$Cohort = "MCI_C"}
  }}


```

#Computing final slopes for each individual (adding back fixed effects)
```{r}

#Adding back fixed effects to random effects to get final slope for each individual
ent_rand_eff$Age = ent_rand_eff$Age + Ent_rnd_slope_int@beta[2]  
ent_rand_eff$`(Intercept)`= ent_rand_eff$`(Intercept)`+ Ent_rnd_slope_int@beta[1]
ent_rand_eff$Coh_slope = Ent_rnd_slope_int@beta[3]
ent_rand_eff$Interaction = Ent_rnd_slope_int@beta[4]

```

#Plotting Results:

```{r}

#fitted values
MCI$Entorhinal_fit = fitted(Ent_rnd_slope_int) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_Entorhinal_Coeff <- ent_rand_eff[ent_rand_eff$Cohort == "MCI_NC",]
MCI_C_Entorhinal_Coeff <- ent_rand_eff[ent_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (age) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Entorhinal_MCI_C_slope = mean(MCI_C_Entorhinal_Coeff$Age, na.rm = TRUE)
Entorhinal_MCI_NC_slope = mean(MCI_NC_Entorhinal_Coeff$Age, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Entorhinal_MCI_C_intercept = mean(MCI_C_Entorhinal_Coeff$`(Intercept)`, na.rm = TRUE)
Entorhinal_MCI_NC_intercept = mean(MCI_NC_Entorhinal_Coeff$`(Intercept)`, na.rm = TRUE)

#Add B2 (Cohort) to MCI_C intercept (as MCI_C is coded 1 and MCI_NC is coded 0)
Entorhinal_MCI_C_intercept = Entorhinal_MCI_C_intercept +  Ent_rnd_slope_int@beta[3]

#Add B3 (Interaction) for MCI_C slope (as MCI_C is coded 1 and MCI_NC is coded 0)
Entorhinal_MCI_C_slope = Entorhinal_MCI_C_slope + Ent_rnd_slope_int@beta[4]

ggplot(MCI, aes(x=Age, y=Entorhinal_fit, group = RID, color = Cohort))+geom_line(size=1.5) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age, y=Entorhinal_icv, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 2) + 
  geom_abline(slope = Entorhinal_MCI_C_slope, intercept = Entorhinal_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = Entorhinal_MCI_NC_slope, intercept = Entorhinal_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject") + ylab("Standardized Entorhinal Volume (SD Units)") + ggtitle("Changes in Entorhinal Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5))  
```

T-tests for initial visits

```{r}
Hip =  data.frame(MCI$RID, MCI$Visit,MCI$Hippocampus_icv,MCI$Cohort)

Hip_1 = Hip[MCI$Visit==0,]
Hip_1_C = Hip_1[Hip_1$MCI.Cohort=="MCI_C",]
Hip_1_NC = Hip_1[Hip_1$MCI.Cohort=="MCI_NC",]
t.test(Hip_1_NC$MCI.Hippocampus_icv,Hip_1_C$MCI.Hippocampus_icv)

Ent =  data.frame(MCI$RID, MCI$Visit,MCI$Entorhinal_icv,MCI$Cohort)

Ent_1 = Ent[MCI$Visit==0,]
Ent_1_C = Ent_1[Ent_1$MCI.Cohort=="MCI_C",]
Ent_1_NC = Ent_1[Ent_1$MCI.Cohort=="MCI_NC",]
t.test(Ent_1_NC$MCI.Entorhinal_icv,Ent_1_C$MCI.Entorhinal_icv)
```

#FUSIFORM Gyrus Volume MODEL
```{r}
# Random Slopes + Random Intercept of age for each subject

Fusif_rnd_slope_int <- lmer(Fusiform_icv ~ Age + Cohort + Age : Cohort + (1 + Age|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = FALSE, na.action = na.exclude)
summary(Fusif_rnd_slope_int)

```


```{r}

#Stoing individual coefficients:

#Coefficients
Fusif_rand_eff <- ranef(Fusif_rnd_slope_int)[["RID"]]
Fusif_rand_eff <- cbind(RID=rownames(Fusif_rand_eff),Fusif_rand_eff)
#rownames(Fusif_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

Fusif_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(Fusif_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (Fusif_rand_eff[i,]$RID == MCI_C_ids[j])
    {Fusif_rand_eff[i,]$Cohort = "MCI_C"}
  }}

#Computing final slopes for each individual (adding back fixed effects)


#Adding back fixed effects to random effects to get final slope for each individual
Fusif_rand_eff$Age = Fusif_rand_eff$Age + Fusif_rnd_slope_int@beta[2]  
Fusif_rand_eff$`(Intercept)`= Fusif_rand_eff$`(Intercept)`+ Fusif_rnd_slope_int@beta[1]
Fusif_rand_eff$Coh_slope = Fusif_rnd_slope_int@beta[3]
Fusif_rand_eff$Interaction = Fusif_rnd_slope_int@beta[4]


#Plotting Fusiform rates of change
#fitted values
MCI$Fusiform_fit = fitted(Fusif_rnd_slope_int) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_Fusiform_Coeff <- Fusif_rand_eff[Fusif_rand_eff$Cohort == "MCI_NC",]
MCI_C_Fusiform_Coeff <- Fusif_rand_eff[Fusif_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (age) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Fusiform_MCI_C_slope = mean(MCI_C_Fusiform_Coeff$Age, na.rm = TRUE)
Fusiform_MCI_NC_slope = mean(MCI_NC_Fusiform_Coeff$Age, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Fusiform_MCI_C_intercept = mean(MCI_C_Fusiform_Coeff$`(Intercept)`, na.rm = TRUE)
Fusiform_MCI_NC_intercept = mean(MCI_NC_Fusiform_Coeff$`(Intercept)`, na.rm = TRUE)

#Add B2 (Cohort) to MCI_C intercept (as MCI_C is coded 1 and MCI_NC is coded 0)
Fusiform_MCI_C_intercept = Fusiform_MCI_C_intercept +  Fusif_rnd_slope_int@beta[3]

#Add B3 (Interaction) for MCI_C slope (as MCI_C is coded 1 and MCI_NC is coded 0)
Fusiform_MCI_C_slope = Fusiform_MCI_C_slope + Fusif_rnd_slope_int@beta[4]

```
 
#Fusiform Plot - WEAKER age cohort interaction than entorhinal? Still significant
```{r}
ggplot(MCI, aes(x=Age, y=Fusiform_fit, group = RID, color = Cohort))+geom_line(size=1.5) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age, y=Fusiform_icv, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 2) + 
  geom_abline(slope = Fusiform_MCI_C_slope, intercept = Fusiform_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = Fusiform_MCI_NC_slope, intercept = Fusiform_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject") + ylab("Standardized Fusiform Volume (SD units)") + ggtitle("Changes in Fusiform Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5)) 
```

#VENTRICLES MODEL

```{r}
# Random Slopes + Random Intercept of age for each subject

Ven_rnd_slope_int <- lmer(Ventricles_icv ~ Age + Cohort + Age : Cohort + (1 + Age|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = FALSE, na.action = na.exclude)
summary(Ven_rnd_slope_int)

```

```{r}
#Stoing individual coefficients:

#Coefficients
Ven_rand_eff <- ranef(Ven_rnd_slope_int)[["RID"]]
Ven_rand_eff <- cbind(RID=rownames(Ven_rand_eff),Ven_rand_eff)
#rownames(Ven_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

Ven_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(Ven_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (Ven_rand_eff[i,]$RID == MCI_C_ids[j])
    {Ven_rand_eff[i,]$Cohort = "MCI_C"}
  }}

#Computing final slopes for each individual (adding back fixed effects)


#Adding back fixed effects to random effects to get final slope for each individual
Ven_rand_eff$Age = Ven_rand_eff$Age + Ven_rnd_slope_int@beta[2]  
Ven_rand_eff$`(Intercept)`= Ven_rand_eff$`(Intercept)`+ Ven_rnd_slope_int@beta[1]
Ven_rand_eff$Coh_slope = Ven_rnd_slope_int@beta[3]
Ven_rand_eff$Interaction = Ven_rnd_slope_int@beta[4]


#Plotting Ventricles rates of change
#fitted values
MCI$Ventricles_fit = fitted(Ven_rnd_slope_int) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_Ventricles_Coeff <- Ven_rand_eff[Ven_rand_eff$Cohort == "MCI_NC",]
MCI_C_Ventricles_Coeff <- Ven_rand_eff[Ven_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (age) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Ventricles_MCI_C_slope = mean(MCI_C_Ventricles_Coeff$Age, na.rm = TRUE)
Ventricles_MCI_NC_slope = mean(MCI_NC_Ventricles_Coeff$Age, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Ventricles_MCI_C_intercept = mean(MCI_C_Ventricles_Coeff$`(Intercept)`, na.rm = TRUE)
Ventricles_MCI_NC_intercept = mean(MCI_NC_Ventricles_Coeff$`(Intercept)`, na.rm = TRUE)

#Add B2 (Cohort) to MCI_C intercept (as MCI_C is coded 1 and MCI_NC is coded 0)
Ventricles_MCI_C_intercept = Ventricles_MCI_C_intercept +  Ven_rnd_slope_int@beta[3]

#Add B3 (Interaction) for MCI_C slope (as MCI_C is coded 1 and MCI_NC is coded 0)
Ventricles_MCI_C_slope = Ventricles_MCI_C_slope + Ven_rnd_slope_int@beta[4]


```

#VENTRICLES PLOT - Pretty Strong age-cohort interaction 

```{r}
ggplot(MCI, aes(x=Age, y=Ventricles_fit, group = RID, color = Cohort))+geom_line(size=1.5) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age, y=Ventricles_icv, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 2) + 
  geom_abline(slope = Ventricles_MCI_C_slope, intercept = Ventricles_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = Ventricles_MCI_NC_slope, intercept = Ventricles_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject") + ylab("Standardized Ventricles Volume (SD units)") + ggtitle("Changes in Ventricles Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5))  
```

```{r}
# Random Slopes + Random Intercept of age for each subject

Occip_rnd_slope_int <- lmer(Occipital_icv ~ Age + Cohort + Age : Cohort + (1 + Age|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = FALSE, na.action = na.exclude)
summary(Occip_rnd_slope_int)

```


```{r}

#Storing individual coefficients:

#Coefficients
Occip_rand_eff <- ranef(Occip_rnd_slope_int)[["RID"]]
Occip_rand_eff <- cbind(RID=rownames(Occip_rand_eff),Occip_rand_eff)
#rownames(Occip_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

Occip_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(Occip_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (Occip_rand_eff[i,]$RID == MCI_C_ids[j])
    {Occip_rand_eff[i,]$Cohort = "MCI_C"}
  }}

#Computing final slopes for each individual (adding back fixed effects)


#Adding back fixed effects to random effects to get final slope for each individual
Occip_rand_eff$Age = Occip_rand_eff$Age + Occip_rnd_slope_int@beta[2]  
Occip_rand_eff$`(Intercept)`= Occip_rand_eff$`(Intercept)`+ Occip_rnd_slope_int@beta[1]
Occip_rand_eff$Coh_slope = Occip_rnd_slope_int@beta[3]
Occip_rand_eff$Interaction = Occip_rnd_slope_int@beta[4]


#Plotting Occipital rates of change
#fitted values
MCI$Occipital_fit = fitted(Occip_rnd_slope_int) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_Occipital_Coeff <- Occip_rand_eff[Occip_rand_eff$Cohort == "MCI_NC",]
MCI_C_Occipital_Coeff <- Occip_rand_eff[Occip_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (age) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Occipital_MCI_C_slope = mean(MCI_C_Occipital_Coeff$Age, na.rm = TRUE)
Occipital_MCI_NC_slope = mean(MCI_NC_Occipital_Coeff$Age, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
Occipital_MCI_C_intercept = mean(MCI_C_Occipital_Coeff$`(Intercept)`, na.rm = TRUE)
Occipital_MCI_NC_intercept = mean(MCI_NC_Occipital_Coeff$`(Intercept)`, na.rm = TRUE)

#Add B2 (Cohort) to MCI_C intercept (as MCI_C is coded 1 and MCI_NC is coded 0)
Occipital_MCI_C_intercept = Occipital_MCI_C_intercept +  Occip_rnd_slope_int@beta[3]

#Add B3 (Interaction) for MCI_C slope (as MCI_C is coded 1 and MCI_NC is coded 0)
Occipital_MCI_C_slope = Occipital_MCI_C_slope + Occip_rnd_slope_int@beta[4]

```

#Occipital Plot - WEAKER age cohort interaction than entorhinal? Still significant
```{r}
ggplot(MCI, aes(x=Age, y=Occipital_fit, group = RID, color = Cohort))+geom_line(size=1.5) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age, y=Occipital_icv, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 2) + 
  geom_abline(slope = Occipital_MCI_C_slope, intercept = Occipital_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = Occipital_MCI_NC_slope, intercept = Occipital_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject") + ylab("Standardized Occipital Volume (SD units)") + ggtitle("Changes in Occipital Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5)) 
```


```{r}
# Random Slopes + Random Intercept of age for each subject

WB_rnd_slope_int <- lmer(WholeBrain_icv ~ Age + Cohort + Age : Cohort + (1 + Age|RID), data = MCI, control = lmerControl(optimizer ="Nelder_Mead"), REML = FALSE, na.action = na.exclude)
summary(WB_rnd_slope_int)

```

```{r}

#Storing individual coefficients:

#Coefficients
WB_rand_eff <- ranef(WB_rnd_slope_int)[["RID"]]
WB_rand_eff <- cbind(RID=rownames(WB_rand_eff),WB_rand_eff)
#rownames(WB_rand_eff)

MCI_NC_ids = unique(MCI$RID[MCI$Cohort == "MCI_NC"])
MCI_C_ids = unique(MCI$RID[MCI$Cohort == "MCI_C"])

WB_rand_eff$Cohort = "MCI_NC"

for (i in 1:nrow(WB_rand_eff)){
  for (j in 1:length(MCI_C_ids)){
    if (WB_rand_eff[i,]$RID == MCI_C_ids[j])
    {WB_rand_eff[i,]$Cohort = "MCI_C"}
  }}

#Computing final slopes for each individual (adding back fixed effects)

#Adding back fixed effects to random effects to get final slope for each individual
WB_rand_eff$Age = WB_rand_eff$Age + WB_rnd_slope_int@beta[2]  
WB_rand_eff$`(Intercept)`= WB_rand_eff$`(Intercept)`+ WB_rnd_slope_int@beta[1]
WB_rand_eff$Coh_slope = WB_rnd_slope_int@beta[3]
WB_rand_eff$Interaction = WB_rnd_slope_int@beta[4]


#Plotting WholeBrain rates of change
#fitted values
MCI$WholeBrain_fit = fitted(WB_rnd_slope_int) #use na.exclude in the model, otherwise size of orignial data and fitted data will be different

#Dividing up coefficients by cohort
MCI_NC_WholeBrain_Coeff <- WB_rand_eff[WB_rand_eff$Cohort == "MCI_NC",]
MCI_C_WholeBrain_Coeff <- WB_rand_eff[WB_rand_eff$Cohort == "MCI_C",]

#Computing the mean B1 (age) coefficient for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
WholeBrain_MCI_C_slope = mean(MCI_C_WholeBrain_Coeff$Age, na.rm = TRUE)
WholeBrain_MCI_NC_slope = mean(MCI_NC_WholeBrain_Coeff$Age, na.rm = TRUE)

#Computing the mean B0 (intercept) for MCI_C and MCI_NC separately (EQUAL TO FIXED EFFECT)
WholeBrain_MCI_C_intercept = mean(MCI_C_WholeBrain_Coeff$`(Intercept)`, na.rm = TRUE)
WholeBrain_MCI_NC_intercept = mean(MCI_NC_WholeBrain_Coeff$`(Intercept)`, na.rm = TRUE)

#Add B2 (Cohort) to MCI_C intercept (as MCI_C is coded 1 and MCI_NC is coded 0)
WholeBrain_MCI_C_intercept = WholeBrain_MCI_C_intercept +  WB_rnd_slope_int@beta[3]

#Add B3 (Interaction) for MCI_C slope (as MCI_C is coded 1 and MCI_NC is coded 0)
WholeBrain_MCI_C_slope = WholeBrain_MCI_C_slope + WB_rnd_slope_int@beta[4]

```

#WholeBrain Plot - WEAKER age cohort interaction than entorhinal? Still significant
```{r}
ggplot(MCI, aes(x=Age, y=WholeBrain_fit, group = RID, color = Cohort))+geom_line(size=1.5) + 
  geom_point(colour = ifelse(MCI$Cohort == "MCI_C", 'darkslategrey','red'),size = 1) +  
  geom_point(aes(x=Age, y=WholeBrain_icv, group = RID, color= Cohort), size = 1.5, stroke = 0, colour = ifelse(MCI$Cohort == "MCI_C", 'darkblue','firebrick4'),size = 2) + 
  geom_abline(slope = WholeBrain_MCI_C_slope, intercept = WholeBrain_MCI_C_intercept, colour = 'darkblue',size = 1.5) +
  geom_abline(slope = WholeBrain_MCI_NC_slope, intercept = WholeBrain_MCI_NC_intercept, colour = 'firebrick4',size = 1.5) +
  xlab("Age of Subject") + ylab("Standardized WholeBrain Volume (SD units)") + ggtitle("Changes in WholeBrain Volume with Age")+
  theme(plot.title = element_text(hjust = 0.5)) 
```